### 2018-06-09 BATexFiles/Draft.tex
removing some files, changing some others

\documentclass{report}
\title{Title}
\author{Thomas Verweyen}
\date{\today}
\begin{document}
\maketitle
\renewcommand*\contentsname{Summary}
\tableofcontents
\chapter{Introduction}
What do i do? I'm working on the Sokoban problem and trying to do my best to find a fast and easy solution to it. I'm not really working on any Sokobanlevel but just the ones meeting some criteria. I'm picking qualitative and quantitative characteristics and deciding on some values to examine. My encodings are supposed to work better on some of the instances and worse on others. What rules describe those fluctuations? Is there a trend when comparing runtimes to my qualitative and quantitative characteristics? E.g. you would expect the runtime to go up when the instance's size grows or when the amount of boxes in the level grows.
In this thesis I will work on optimizing ASP encodings for the game Sokoban towards different instances.
\chapter{What is Sokoban}
Sokoban is a logistics game, where the player controls a character in a warehouse. The goal of the game is to move all crates in the warehouse on one of the target fields.
\chapter{State of the Art}
- mention other works on Sokoban
- what tests?
- what results?
- whats the connection?
\chapter{My Work}
what did i do?
what is my baseline?
what changes from encoding to encoding
which of the encodings works best in which kind of instance
how do i evaluate that
show my experiments and their results
	\section{Encodings}
	\section{Evaluation}
	\section{Experiments and Results}
\chapter{Discussion of Results}
compare chapter State of the Art with chapter My Work
\end{document}

### 2018-06-09 Stuff/Parser.py

@@ -1,0 +1,211 @@
+def read_map():
+    f = open('newinstance02.lp', 'r')
+    sr = f.read()
+    # f = open('atlist','r')
+
+    mapRepresentation = []
+    playerPosition = [0, 0]
+    boxPos = []
+
+    # find out what the dimensions of the map are ...
+    n = get_max(sr)
+    # ... and creating the boolean representation of it
+    for i in range(n[1]):
+        helplist = []
+        for j in range(n[0]):
+            helplist += [False]
+        mapRepresentation += [helplist]
+    boxPos = []
+    for k in range(n[2]):
+        boxPos += [[0, 0]]
+    targetPos = []
+    for l in range(n[2]):
+        targetPos += [[0, 0]]
+
+    # calling readMap and showMap
+    linelist = sr.splitlines(True)
+    indexTargets = 0
+    for line in linelist:
+        # read out fields, by checking for the init(field((a;b..c),d)) syntax
+        if line[5:10] == "field":
+            s = line[11:]
+            a = ""
+            b = ""
+            lr = True
+            la = []
+            lb = []
+            while len(s) > 0:
+                if (ord(s[0]) in range(48, 58)) or (s[0] == ';'):
+                    if lr:
+                        a = a + s[0]
+                    else:
+                        b = b + s[0]
+                    s = s[1:]
+                elif s[0:2] == "..":
+                    if lr:
+                        a = a + s[0:2]
+                    else:
+                        b = b + s[0:2]
+                    s = s[2:]
+                elif s[0] == ',':
+                    lr = False
+                    l = lb
+                    la += finish(a)
+                    a = ""
+                    lb += finish(b)
+                    b = ""
+                    s = s[1:]
+                elif s[0] == '.':
+                    l = lb
+                    la += finish(a)
+                    a = ""
+                    lb += finish(b)
+                    b = ""
+                    s = s[1:]
+                else:
+                    l = lb
+                    la += finish(a)
+                    a = ""
+                    lb += finish(b)
+                    b = ""
+                    s = s[1:]
+            for i in lb:
+                for j in la:
+                    mapRepresentation[i - 1][j - 1] = True
+        elif line[5:7] == "at":
+            s = line[8:]
+            a = ""
+            lr = True
+            u = 0
+            stringlist = s.split(",")
[…]

### 2018-06-09 NewGenerator.py
asd

@@ -1,392 +1,0 @@
-from math import floor, sqrt
-from itertools import chain
-from copy import deepcopy
-
-
-# coding=utf-8
-def main(dim_x, dim_y, box_number):
-    level_index = 0
-    fields = []
-    for i in range(0, dim_x):
-        x = []
-        for j in range(0, dim_y):
-            x += [False]
-        fields += [x]
-    list_of_basic_levels = [fields]
-    # try to make ALL the levels
-    # the way it's done here is by creating the base of only walls ([[False,False,..],[False,..],..]) and
-    # changing it in all possible ways (given by the get_options 1-4)
-    while len(list_of_basic_levels) > 0:
-        this_level = list_of_basic_levels[0]
-        list_of_basic_levels = list_of_basic_levels[1:]
-        # place fields
-        if count_fields(this_level) < (dim_x * dim_y * 3 / 4):
-            if count_fields(this_level) < (floor(dim_x * dim_y * 4 / 10)):
-                the_blocks = get2x2options(dim_x, dim_y, this_level)
-                for block in the_blocks:
-                    list_of_basic_levels += [make2x2block(block[0], block[1], this_level)]
-            else:
-                # get all options into one big list
-                for option in chain(get_options1(dim_x, dim_y, this_level), get_options2(dim_x, dim_y, this_level),
-                                    get_options3(dim_x, dim_y, this_level), get_options4(dim_x, dim_y, this_level)):
-                    list_of_basic_levels += [change_fields_by(this_level, option)]
-        # place boxes and targets
-        else:
-            boxes_placed = 0
-            targets_placed = 0
-            list_of_levels_w_boxes = [this_level]
-            # index 0 is player, 1 is box, 2 is target, 3 is player-target, 4 is box-target
-            # TODO the program places only one box regardless of the given box_number -- FIX THAT
-            while boxes_placed < box_number:
-                my_iterator_list = deepcopy(list_of_levels_w_boxes)
-                list_of_levels_w_boxes = []
-                for my_list_ele in my_iterator_list:
-                    box_options = get_box_options(dim_x, dim_y, my_list_ele)
-                    for opts in box_options:
[…]

### 2018-06-09 Parser.py

@@ -1,211 +1,0 @@
-def read_map():
-    f = open('newinstance02.lp', 'r')
-    sr = f.read()
-    # f = open('atlist','r')
-
-    mapRepresentation = []
-    playerPosition = [0, 0]
-    boxPos = []
-
-    # find out what the dimensions of the map are ...
-    n = get_max(sr)
-    # ... and creating the boolean representation of it
-    for i in range(n[1]):
-        helplist = []
-        for j in range(n[0]):
-            helplist += [False]
-        mapRepresentation += [helplist]
-    boxPos = []
-    for k in range(n[2]):
-        boxPos += [[0, 0]]
-    targetPos = []
-    for l in range(n[2]):
-        targetPos += [[0, 0]]
-
-    # calling readMap and showMap
-    linelist = sr.splitlines(True)
-    indexTargets = 0
-    for line in linelist:
-        # read out fields, by checking for the init(field((a;b..c),d)) syntax
-        if line[5:10] == "field":
-            s = line[11:]
-            a = ""
-            b = ""
-            lr = True
-            la = []
-            lb = []
-            while len(s) > 0:
-                if (ord(s[0]) in range(48, 58)) or (s[0] == ';'):
-                    if lr:
-                        a = a + s[0]
-                    else:
-                        b = b + s[0]
-                    s = s[1:]
-                elif s[0:2] == "..":
-                    if lr:
-                        a = a + s[0:2]
-                    else:
-                        b = b + s[0:2]
-                    s = s[2:]
-                elif s[0] == ',':
-                    lr = False
-                    l = lb
-                    la += finish(a)
-                    a = ""
-                    lb += finish(b)
-                    b = ""
-                    s = s[1:]
-                elif s[0] == '.':
-                    l = lb
-                    la += finish(a)
-                    a = ""
-                    lb += finish(b)
-                    b = ""
-                    s = s[1:]
-                else:
-                    l = lb
-                    la += finish(a)
-                    a = ""
-                    lb += finish(b)
-                    b = ""
-                    s = s[1:]
-            for i in lb:
-                for j in la:
-                    mapRepresentation[i - 1][j - 1] = True
-        elif line[5:7] == "at":
-            s = line[8:]
-            a = ""
-            lr = True
-            u = 0
-            stringlist = s.split(",")
[…]

### 2018-06-09 SokoShow.py

@@ -1,259 +1,0 @@
-import sys
-import os
-from PyQt5 import QtGui, QtWidgets
-from PyQt5 import QtCore
-
-
-class SokoShow(QtWidgets.QApplication):
-    def main(self):
-        indT = 0
-
-        def readMap(mapping):
-            linelist = mapping.splitlines(True)
-            indexTargets = 0
-            for line in linelist:
-                # read out fields, by checking for the init(field((a;b..c),d)) syntax
-                if line[5:10] == "field":
-                    s = line[11:]
-                    a = ""
-                    b = ""
-                    lr = True
-                    la = []
-                    lb = []
-                    while len(s) > 0:
-                        if (ord(s[0]) in range(48, 58)) or (s[0] == ';'):
-                            if lr:
-                                a = a + s[0]
-                            else:
-                                b = b + s[0]
-                            s = s[1:]
-                        elif s[0:2] == "..":
-                            if lr:
-                                a = a + s[0:2]
-                            else:
-                                b = b + s[0:2]
-                            s = s[2:]
-                        elif s[0] == ',':
-                            lr = False
-                            l = lb
-                            la += finish(a)
-                            a = ""
-                            lb += finish(b)
-                            b = ""
-                            s = s[1:]
-                        elif s[0] == '.':
-                            l = lb
-                            la += finish(a)
-                            a = ""
-                            lb += finish(b)
-                            b = ""
-                            s = s[1:]
-                        else:
-                            l = lb
-                            la += finish(a)
-                            a = ""
-                            lb += finish(b)
-                            b = ""
-                            s = s[1:]
-                    for i in lb:
-                        for j in la:
-                            mapRepresentation[i - 1][j - 1] = True
-                elif line[5:7] == "at":
-                    s = line[8:]
-                    a = ""
-                    lr = True
-                    u = 0
-                    stringlist = s.split(",")
-                    stringlist[2] = stringlist[2].translate(None, ',).\n')
-                    if int(stringlist[0]) == 0:
-                        playerPosition = [int(stringlist[1]), int(stringlist[2])]
-                    else:
-                        boxPos[int(stringlist[0]) - 1] = [int(stringlist[1]), int(stringlist[2])]
[…]

### 2018-06-09 Visualizer.py

@@ -1,127 +1,0 @@
-import sys
-import os
-from Parser import read_map, get_max
-from PyQt5 import QtGui, QtWidgets
-
-step_index = 0
-max_step = 1
-
-
-class Visualizer(QtWidgets.QApplication):
-
-    def __init__(self):
-        QtWidgets.QApplication.__init__(self, sys.argv)
-        w = QtWidgets.QWidget()
-        w.setWindowTitle('')
-        my_grid = QtWidgets.QGridLayout()
-        w.setLayout(my_grid)
-
-        f = open('newinstance03.lp', 'r')
-        sr = f.read()
-        n = get_max(sr)
-        self.redraw(my_grid, w)
-        # w.resize(boxpic.width(),boxpic.height())
-        button0 = QtWidgets.QPushButton("Previous Step", w)
-        button1 = QtWidgets.QPushButton("Next Step", w)
-        button2 = QtWidgets.QPushButton("From Beginning", w)
-
-        def button_back():
-            global step_index
-            if step_index > 0:
-                step_index -= 1
-                print(step_index)
-                self.redraw(my_grid, w)
-
-        def button_forward():
-            global step_index
-            if step_index < max_step:
-                step_index += 1
-                print(step_index)
-                self.redraw(my_grid, w)
-
-        def button_beginning():
-            global step_index
-            step_index = 0
-            print(step_index)
-            self.redraw(my_grid, w)
-
-        button0.clicked.connect(button_back)
-        button1.clicked.connect(button_forward)
-        button2.clicked.connect(button_beginning)
-        my_grid.addWidget(button0, n[1] - 3, n[0])
-        my_grid.addWidget(button1, n[1] - 2, n[0])
-        my_grid.addWidget(button2, n[1] - 1, n[0])
-        w.setGeometry(50,50,500,400)
-        w.show()
-        sys.exit(self.exec_())
-
-    def redraw(self, grid, widget):
-        mapping, boxes, targets, player, at_list, max_t = read_map()
-        global max_step
-        if max_t > max_step:
-            max_step = max_t
-        box = QtGui.QPixmap(os.getcwd() + '/Img/Box.png')
-        empty = QtGui.QPixmap(os.getcwd() + '/Img/Empty.png')
-        guy = QtGui.QPixmap(os.getcwd() + '/Img/Guy.png')
-        wall = QtGui.QPixmap(os.getcwd() + '/Img/Wall.png')
[…]

### 2018-07-16 BATexFiles/Draft.tex
Added a lot of text to the texfile

@@
 \date{\today}
+\usepackage{listings}
+\usepackage{color}
+\lstset{numbers=left,
+literate = {-}{-}1,
+basicstyle=\footnotesize,
+numbersep=5pt, backgroundcolor=\color{white},
+captionpos=b,keywordstyle=\color{blue},
+commentstyle=\color{green}}
 \begin{document}
@@
 \tableofcontents
+\section*{Abstract}
+Sokoban approach with ASP planning -$>$ Plain, Inc, Optimization (?)
 \chapter{Introduction}
@@
 I'm working on the Sokoban problem and trying to do my best to find a fast and easy solution to it.
-I'm not really working on any Sokobanlevel but just the ones meeting some criteria.
+I'm not working on any level but just the ones meeting some criteria.
 I'm picking qualitative and quantitative characteristics and deciding on some values to examine.
@@
 \chapter{What is Sokoban}
-Sokoban is a logistics game, where the player controls a character in a warehouse. The goal of the game is to move all crates in the warehouse on one of the target fields.
+Sokoban is a logistics game, where the player controls a character in a warehouse.
+The goal of the game is to move all crates in the warehouse on one of the target fields.
+The rules state that crates cannot be stacked on one field and crates cannot be pushed if there is a crate or a wall blocking the way.
+Each player move is a step into one of the four directions in a two-dimensional grid.
+If the player would collide with a crate, the crate is pushed into the direction of the player-movement.
+A level is usually satisfied when each crate is pushed on a target; also we can conclude that each target would be covered with a crate, because the amount of targets and crates is equal.
 \chapter{State of the Art}
@@
 \section{Encodings}
+\subsection{Plain Encoding}
+\lstinputlisting{../Encodings/SokobanPlainEncoding.lp}
+The first try to construct a naive encoding to solve sokoban-instances yielded a solution that works in a common way in ASP planning.
+In line 4 we initialize the information from the level instance with the time 0.
+In line 7 we state the there is one move at minimum and at maximum for each timestep.
+This makes the ASP planner guess the sequence of moves and infer every successive state from the moves it picks.
+In lines 9-13 we calculate the successor state from the move picked in line 7.
+Line 9 calculates the new player-position, line 11 calculates the new position of the box in the players way and line 13 calculates the new position of boxes not touched by the player.
+Additionally we need some integrity constraints to prevent illegal states.
+The integrity constraint in line 15 ensures that at no time a box or player has a position outside of the playing field.
+The integrity constraint in line 17 ensures that at no time boxes or the player share one field.
+In the lines 19-21 we define our win-condition to be that all targets in the instance have to be covered by boxes.
+The Plain Encoding has the predicates, time/1, dir/2, at/4, move/3, uncovered/3, goal/3
+The encoding works in a common way in ASP planning, by guessing the sequence of actions and inferring the successor states by effect axioms.
+There is a time/1 predicate for every timestep and a at/4 predicate for every box and the player at every timestep.
+\subsection{Incremental Encoding}
+\lstinputlisting{../Encodings/SokobanPlainEncodingInc.lp}
+The incremental encoding is similar to the plain encoding.
+The difference is the partition of the encoding into three parts: the \#program base in lines 3-5 which is independent of the step parameter t, the \#program step in lines 7-19 which is the cumulative part, collecting knowledge and the \#program check in lines 21-25 which is specific for each value of t.
+This allows for gradually processing each time step and accumulating knowledge that persists even if the current step is not satisfiable.
+\subsection{Advanced Encoding}
+\lstinputlisting{../Encodings/SokobanPlainEncodingOpti.lp}
 \section{Evaluation}
@@
 \chapter{Discussion of Results}
-compare chapter State of the Art with chapter My Work
+compare chapter State of the Art with chapter My Work \cite{LifschitzASP}
+
+\begin{thebibliography}{}
+\bibitem{LifschitzASP}
+Vladimir Lifschitz
+\textit{Answer Set Planning}
+ICLP, 1999.
+
+\bibitem{Lifschitz2002AnswerSP}
+Vladimir Lifschitz
+\textit{Answer set programming and plan generation}
+Artificial Intelligence, 2002, Vol. 138, p.39-54.
+
+\bibitem{LifschitzWhatisASP}
+Vladimir Lifschitz
+\textit{What is Answer Set Programming?}
+AAAI, 2008.
+
+\bibitem{Gebser2013RicochetRA}
+Martin Gebser, Holger Jost, Roland Kaminski, Philipp Obermeier, Orkunt Sabuncu, Torsten Schaub, Marius Thomas Lindauer
+\textit{Ricochet Robots: A Transverse ASP Benchmark}
+LPMNR, 2013.
+
+\bibitem{BoteaHeuristicsVsPlanning}
+Adi Botea, Martin Müller, Jonathan Schaeffer
+\textit{Using Abstraction for Planning in Sokoban}
+Computers and Games, 2002.
+
+\bibitem{Frisch2006PropagationAF}
+Alan M. Frisch, Brahim Hnich, Zeynep Kiziltan, Ian Miguel, Toby Walsh
+\textit{Propagation algorithms for lexicographic ordering constraints}
+Artificial Intelligence, 2006, Vol. 170, p.803-834.
+
+\bibitem{Dor1999SOKOBANAO}
+Dorit Dor, Uri Zwick
+\textit{SOKOBAN and other motion planning problems}
+Computational Geometry, 1999, Vol. 13, p.215-228.
+
+\bibitem{Froleyks2017UsingAA}
+Nils Christian Froleyks and Tom{\'a}{\^s} Balyo
+\textit{Using an Algorithm Portfolio to Solve Sokoban}
+SOCS, 2017.
+
+\bibitem{PSpaceComplete}
+Joseph C. Culberson
+\textit{Sokoban is PSPACE-complete}
+1997.
+
+\bibitem{SokoRelevanceCuts}
+Andreas Junghanns, Jonathan Schaeffer
+\textit{Sokoban: Improving the Search with Relevance Cuts}
+Journal of Theoretical Computing Science, 1999, Vol. 252, p.1-2.
+
+\bibitem{SymmetryBreaking}
+Brahim Hnich, Zeynep Kiziltan, Toby Walsh
+\textit{Combining Symmetry Breaking with Other Constraints: Lexicographic Ordering with Sums}
+ISAIM, 2004.
+\end{thebibliography}
+
 \end{document}

### 2018-07-16 Stuff/Parser.py

@@ -1,5 +1,5 @@
 def read_map():
-    f = open('newinstance02.lp', 'r')
+    f = open('InstancesClean/Inst-A+Z20-4Halle.lp', 'r')
     sr = f.read()
     # f = open('atlist','r')
 
@@ -95,11 +95,9 @@ def read_map():
             indexTargets += 1
     if playerPosition == None:
         exit('no player position')
-    f = open('action_plan.txt','r')
+    f = open('Stuff/action_plan.txt','r')
     sr = f.read()
     at_string_arr = sr.split(" ")
-
-    print(at_string_arr)
     # cut line end
     at_string_arr[len(at_string_arr)-1] = at_string_arr[len(at_string_arr)-1][:len(at_string_arr[len(at_string_arr)-1])-1]
 

### 2018-07-16 Stuff/Visualizer.py

@@ -16,7 +16,7 @@ class Visualizer(QtWidgets.QApplication):
         my_grid = QtWidgets.QGridLayout()
         w.setLayout(my_grid)
 
-        f = open('newinstance03.lp', 'r')
+        f = open('InstancesClean/Inst-A+Z20-4Halle.lp', 'r')
         sr = f.read()
         n = get_max(sr)
         self.redraw(my_grid, w)

### 2018-07-16 BATexFiles/Bibliography.tex
added some files that were not correctly tracked

--- 2018-06-09 BATexFiles/Draft.tex
@@
-\documentclass{report}
-\title{Title}
-\author{Thomas Verweyen}
-\date{\today}
-\begin{document}
-\maketitle
-\renewcommand*\contentsname{Summary}
-\tableofcontents
-\chapter{Introduction}
-What do i do?
-I'm working on the Sokoban problem and trying to do my best to find a fast and easy solution to it.
-I'm not really working on any Sokobanlevel but just the ones meeting some criteria.
-I'm picking qualitative and quantitative characteristics and deciding on some values to examine.
-My encodings are supposed to work better on some of the instances and worse on others.
-What rules describe those fluctuations?
-Is there a trend when comparing runtimes to my qualitative and quantitative characteristics?
-E.g.
-you would expect the runtime to go up when the instance's size grows or when the amount of boxes in the level grows.
-In this thesis I will work on optimizing ASP encodings for the game Sokoban towards different instances.
-\chapter{What is Sokoban}
-Sokoban is a logistics game, where the player controls a character in a warehouse. The goal of the game is to move all crates in the warehouse on one of the target fields.
-\chapter{State of the Art}
-- mention other works on Sokoban
-- what tests?
-- what results?
-- whats the connection?
-\chapter{My Work}
-what did i do?
-what is my baseline?
-what changes from encoding to encoding
-which of the encodings works best in which kind of instance
-how do i evaluate that
-show my experiments and their results
-	\section{Encodings}
-	\section{Evaluation}
-	\section{Experiments and Results}
-\chapter{Discussion of Results}
-compare chapter State of the Art with chapter My Work
-\end{document}
+\begin{thebibliography}{}
+\bibitem{LifschitzASP}
+Vladimir Lifschitz
+\textit{Answer Set Planning}
+ICLP, 1999.
+
+\bibitem{Lifschitz2002AnswerSP}
+Vladimir Lifschitz
+\textit{Answer set programming and plan generation}
+Artificial Intelligence, 2002, Vol. 138, p.39-54.
+
+\bibitem{LifschitzWhatisASP}
+Vladimir Lifschitz
+\textit{What is Answer Set Programming?}
+AAAI, 2008.
+
+\bibitem{Gebser2013RicochetRA}
+Martin Gebser, Holger Jost, Roland Kaminski, Philipp Obermeier, Orkunt Sabuncu, Torsten Schaub, Marius Thomas Lindauer
+\textit{Ricochet Robots: A Transverse ASP Benchmark}
+LPMNR, 2013.
+
+\bibitem{BoteaHeuristicsVsPlanning}
+Adi Botea, Martin Müller, Jonathan Schaeffer
+\textit{Using Abstraction for Planning in Sokoban}
+Computers and Games, 2002.
+
+\bibitem{Frisch2006PropagationAF}
+Alan M. Frisch, Brahim Hnich, Zeynep Kiziltan, Ian Miguel, Toby Walsh
+\textit{Propagation algorithms for lexicographic ordering constraints}
+Artificial Intelligence, 2006, Vol. 170, p.803-834.
+
+\bibitem{Dor1999SOKOBANAO}
+Dorit Dor, Uri Zwick
+\textit{SOKOBAN and other motion planning problems}
+Computational Geometry, 1999, Vol.13, p.215-228.
+
+\bibitem{Froleyks2017UsingAA}
+Nils Christian Froleyks and Tom{\'a}s Balyo
+\textit{Using an Algorithm Portfolio to Solve Sokoban}
+SOCS, 2017.
+\end{thebibliography}

### 2018-07-28 BATexFiles/Draft.tex
Refined some wordings in the Tex, tried around with the plain-f+c-split encoding to get it to order lexicographically

@@
 \documentclass{report}
-\title{Title}
+\title{Efficient Planning for Sokoban with Answer Set Programming - A Case Study}
 \author{Thomas Verweyen}
@@
 \section*{Abstract}
-Sokoban approach with ASP planning -$>$ Plain, Inc, Optimization (?)
+Sokoban approach with ASP planning $\rightarrow$ PlainEnc, IncEnc, Optimization (?)
 \chapter{Introduction}
@@
 What rules describe those fluctuations?
+What assumptions can be made?
 Is there a trend when comparing runtimes to my qualitative and quantitative characteristics?
@@
 In this thesis I will work on optimizing ASP encodings for the game Sokoban towards different instances.
+First Ill briefly explain what Sokoban is and how it is played, then Ill give a short introduction on comparable/similar/other scientific studies concerning sokoban.
+Then Ill go over my work, ...
+and finally evaluation and Discussion of Results.
 \chapter{What is Sokoban}
@@
 In line 4 we initialize the information from the level instance with the time 0.
-In line 7 we state the there is one move at minimum and at maximum for each timestep.
-This makes the ASP planner guess the sequence of moves and infer every successive state from the moves it picks.
+In line 7 we state that there is one move at minimum and at maximum for each timestep.
+This makes the ASP planner ground all sequences of moves, pick one for every timestep and infer every successive state from the moves it picks.
 In lines 9-13 we calculate the successor state from the move picked in line 7.
@@
 There is a time/1 predicate for every timestep and a at/4 predicate for every box and the player at every timestep.
+
 \subsection{Incremental Encoding}
@@
 The incremental encoding is similar to the plain encoding.
-The difference is the partition of the encoding into three parts: the \#program base in lines 3-5 which is independent of the step parameter t, the \#program step in lines 7-19 which is the cumulative part, collecting knowledge and the \#program check in lines 21-25 which is specific for each value of t.
+The difference is the partition of the encoding into three parts: the \#program base in lines 3-5 which is independent of the step parameter t, the \#program step in lines 7-19 which is the cumulative part, collecting knowledge and the \#program check in lines 21-25 which is specific for each value of t (STEP ALSO IS THOUGH).
 This allows for gradually processing each time step and accumulating knowledge that persists even if the current step is not satisfiable.
+
 \subsection{Advanced Encoding}

### 2018-08-17 BATexFiles/Draft.tex
Added a lot of little textes and Information

@@
 \usepackage{listings}
+\usepackage{graphicx}
+\graphicspath{{./images/}}
 \usepackage{color}
@@
 \chapter{What is Sokoban}
-Sokoban is a logistics game, where the player controls a character in a warehouse.
+Sokoban is a logistic planning puzzle.
+The player pushes boxes around a warehouse, trying to cover all goal fields with one of the boxes.
+For that the player can move one step at a time into any of the four directions of a two-dimensional grid.
+When the player moves onto a box's position the box is pushed one field into the direction of the move.
+Each box that is not pushed at step T stays at the same position for step T+1.
+There can never be multiple boxes on one field, neither can there be a box where there is no field.
+This is a very difficult planning problem for multiple reasons.
+There are a lot of possibilities to create a deadlock (a gamestate from which no solution can be found), the branching factor is large and the optimal solution can be very long \cite{BoteaHeuristicsVsPlanning}.
+
+%Sokoban is a logistics game, where the player controls a character in a warehouse.
 The goal of the game is to move all crates in the warehouse on one of the target fields.
@@
 \chapter{State of the Art}
+
 - mention other works on Sokoban
@@
 \section{Encodings}
+\subsection{Fact Format}
+\includegraphics{Visualizer}
+\lstinputlisting{../InstancesClean/Inst-A-Z20-6Halle.lp}
 \subsection{Plain Encoding}

### 2018-08-17 Stuff/Parser.py

@@ -1,5 +1,5 @@
 def read_map():
-    f = open('InstancesClean/Inst-A+Z20-4Halle.lp', 'r')
+    f = open('InstancesClean/Inst-A-Z20-6Halle.lp', 'r')
     sr = f.read()
     # f = open('atlist','r')
 

### 2018-08-17 Stuff/Visualizer.py

@@ -16,7 +16,7 @@ class Visualizer(QtWidgets.QApplication):
         my_grid = QtWidgets.QGridLayout()
         w.setLayout(my_grid)
 
-        f = open('InstancesClean/Inst-A+Z20-4Halle.lp', 'r')
+        f = open('InstancesClean/Inst-A-Z20-6Halle.lp', 'r')
         sr = f.read()
         n = get_max(sr)
         self.redraw(my_grid, w)

### 2018-08-19 BATexFiles/Draft.tex
Working on that tex file ... GOGOGOGOGOGO

@@
 Sokoban approach with ASP planning $\rightarrow$ PlainEnc, IncEnc, Optimization (?)
-\chapter{Introduction}
+\section{Introduction}
 What do i do?
@@
 and finally evaluation and Discussion of Results.
-\chapter{What is Sokoban}
-Sokoban is a logistic planning puzzle.
-The player pushes boxes around a warehouse, trying to cover all goal fields with one of the boxes.
-For that the player can move one step at a time into any of the four directions of a two-dimensional grid.
+\section{What is Sokoban}
+\subsection{Background}
+%% ZU KURZ, MIR FEHLT EIN BISSCHEN LORE (SOKOBAN HEISST WARENHAUSMANN, ENTWICKELT VON.., ...) UND EIN BISSCHEN FORMALITÄT (VGL \cite{Froleyks2017UsingAA})
+Sokoban is a logistic planning puzzle game from Japan.
+Sokoban was apparently created by Hiroyuki Imabayashi in 1981, and published by \textit{Thinking Rabbit} in 1982.
+Sokoban means warehouse person in japanese.
+The game has very simple rules but shows to be very complex for humans and for AI.
+LOT OF RESEARCH REGARDING THAT (SO COMPLEX, SO EXEMPLARY, SO RELATABLE TO REAL LIFE AI PROBLEMS, SO EASY TO RESEARCH)
+\subsection{Rules}
+\begin{center}
+\includegraphics[scale=0.2]{SokobanLevel1}
+\newline
+\textbf{\textit{Fig.1: First of the Classical Sokoban Levels}}
+\end{center}
+Figure 1 illustrates the first Sokoban level of the original set of levels.
+The game is played on a board that can be modelled as a two-dimensional grid.
+The squares can either be floor or wall squares.
+Some floor squares are marked as goals and the same number of floor squares contain boxes.
+The player pushes boxes around this modelled warehouse, trying to cover all of the marked goal fields with one of the boxes.
+For that the player can move one step at a time into one of the four directions \textit{up}, \textit{down}, \textit{left} and \textit{right}.
 When the player moves onto a box's position the box is pushed one field into the direction of the move.
-Each box that is not pushed at step T stays at the same position for step T+1.
 There can never be multiple boxes on one field, neither can there be a box where there is no field.
-This is a very difficult planning problem for multiple reasons.
+In many of the levels created to this date, the warehouse has a lot of small passageways and very intricate designs. In this thesis I will look at more open level designs, e.g. rectangular levels.
+Sokoban is a very difficult planning problem for multiple reasons.
 There are a lot of possibilities to create a deadlock (a gamestate from which no solution can be found), the branching factor is large and the optimal solution can be very long \cite{BoteaHeuristicsVsPlanning}.
@@
 A level is usually satisfied when each crate is pushed on a target; also we can conclude that each target would be covered with a crate, because the amount of targets and crates is equal.
-\chapter{State of the Art}
+\section{State of the Art}
 
@@
 - whats the connection?
-\chapter{My Work}
+\section{My Work}
 what did i do?
@@
 show my experiments and their results
-\section{Encodings}
-\subsection{Fact Format}
+\subsection{Encodings}
+\subsubsection{Fact Format}
 \includegraphics{Visualizer}
 \lstinputlisting{../InstancesClean/Inst-A-Z20-6Halle.lp}
-\subsection{Plain Encoding}
+\subsubsection{Plain Encoding}
 \lstinputlisting{../Encodings/SokobanPlainEncoding.lp}
@@
 
-\subsection{Incremental Encoding}
+\subsubsection{Incremental Encoding}
 \lstinputlisting{../Encodings/SokobanPlainEncodingInc.lp}
@@
 
-\subsection{Advanced Encoding}
+\subsubsection{Advanced Encoding}
 \lstinputlisting{../Encodings/SokobanPlainEncodingOpti.lp}
-\section{Evaluation}
-\section{Experiments and Results}
-\chapter{Discussion of Results}
+\subsection{Evaluation}
+\subsection{Experiments and Results}
+\section{Discussion of Results}
 compare chapter State of the Art with chapter My Work \cite{LifschitzASP}

### 2018-08-19 BATexFiles/Draft.tex
added a few lines to the 2nd chapter

@@
-\documentclass{report}
+\documentclass{article}
 \title{Efficient Planning for Sokoban with Answer Set Programming - A Case Study}
@@
 \section{Introduction}
+TODO: REWORK THIS
 What do i do?
@@
 The game has very simple rules but shows to be very complex for humans and for AI.
-LOT OF RESEARCH REGARDING THAT (SO COMPLEX, SO EXEMPLARY, SO RELATABLE TO REAL LIFE AI PROBLEMS, SO EASY TO RESEARCH)
+This makes Sokoban a very interesting subject for research regarding planning problems.
+%LOT OF RESEARCH REGARDING THAT (SO COMPLEX, SO EXEMPLARY, SO RELATABLE TO REAL LIFE AI PROBLEMS, SO EASY TO RESEARCH)
 \subsection{Rules}
@@
 Some floor squares are marked as goals and the same number of floor squares contain boxes.
-The player pushes boxes around this modelled warehouse, trying to cover all of the marked goal fields with one of the boxes.
+The player pushes boxes around this modelled warehouse, trying to cover the marked goal fields (targets) with one of the boxes.
 For that the player can move one step at a time into one of the four directions \textit{up}, \textit{down}, \textit{left} and \textit{right}.
@@
 There can never be multiple boxes on one field, neither can there be a box where there is no field.
+A \textit{solution} to a level is a sequence of moves that leaves a box on each target.
 In many of the levels created to this date, the warehouse has a lot of small passageways and very intricate designs. In this thesis I will look at more open level designs, e.g. rectangular levels.
@@
 \section{State of the Art}
+Gordon Wilfong published 1991 that all motion planning problems with moving obstacles are NP-hard \cite{WilfongNPhard} by reducing 3Sat to such a problem.
+He didnt answer the question whether Sokoban is also PSPACE-complete.
+That was answered by Joseph C Culberson \cite{PSpaceComplete} in 1997.
+He showed that it is possible to emulate a linear bounded automaton by constructing a level.
+The solution to this constructed level, if there is any, will have a length in $\Theta (|w|+t(|w|))$, where $w$ is the input and $t(|w|)$ is the amount of transitions the LBA made.
+This shows PSPACE-completeness
 
-- mention other works on Sokoban
-- what tests?
-- what results?
-- whats the connection?
+Heuristics doesnt seem to do a great deal for sokoban \cite{BoteaHeuristicsVsPlanning}.
+It helped a lot with chess e.g., but doesnt cut it for sokoban.
+Approach the hard domain of sokoban in another way.
+Macromoves a way that really works well and is used in practice, also deadlock detection (WHAT IS A DEADLOCK - stone in corner and no goal -> no solutions) (obviously best if you know the specific level).
+Reference Rolling stone (source not available (?)).
+Furthermore Sokoban can be abstracted into rooms and tunnels, making it into subproblems which should be easier to solve, tunnels for example, are mostly immediately solved.
+Similarly you can also use relevance cutting.
+This means creating a measurement for relevance in between moves and picking moves that are relevant to prior moves (few exceptions in special cases).
+Backwards Approach and LexOrdering.
+Conclusion
+%Sokoban was researched to be PSPACE-complete and NP-hard.
+PSPACE-complete was done by constructing levels that are PSPACE-complete (look at the work again!!).
+Heuristic seems to be of little use, so go for planning instead [botea]
+%Sokoban can be abstracted into subproblems to get more smaller problems, greatly reducing searchspace.
+%This also makes sense in relevance cutting, which looks at relations between boxes to find groups that behave together and cleanly separate those.
+%Alternatively Backwards Approach
+%Or LexOrdering
+%All in all a Plethora of different approaches to crack the problems of sokoban.
+
+%- mention other works on Sokoban
+%- what tests?
+%- what results?
+%- whats the connection?
 \section{My Work}
@@
 ISAIM, 2004.
+
+\bibitem{WilfongNPhard}
+Gordon Wilfong
+\textit{Motion planning in the presence of movable obstacles}
+Annals of Mathematics and Artificial Intelligence, 1991.
 \end{thebibliography}

### 2018-08-20 BATexFiles/Draft.tex
again a few lines changed in the tex

@@
 \section{State of the Art}
-Gordon Wilfong published 1991 that all motion planning problems with moving obstacles are NP-hard \cite{WilfongNPhard} by reducing 3Sat to such a problem.
-He didnt answer the question whether Sokoban is also PSPACE-complete.
+\textbf{Complexity:} Gordon Wilfong published in 1991 that all motion planning problems with moving obstacles are NP-hard \cite{WilfongNPhard}.
+He proved this by reducing $3$-$Sat$ to such a problem.
+The question whether Sokoban is PSPACE-complete remained open until 1997, when Joseph C.
+Culberson published his paper "Sokoban is PSPACE-complete" \cite{PSpaceComplete}.
+He found a transformation algorithm to transform an LBA into a level of Sokoban, with the solution to the level having a number of steps in $\Theta (|w|+t(|w|))$, where $w$ is the input and $t(|w|)$ is the amount of transitions the LBA made.
+%This shows PSPACE-hardness and Sokoban is known to be in PSPACE, so Sokoban is PSPACE-complete.
+\newline%He didnt answer the question whether Sokoban is also PSPACE-complete.
 That was answered by Joseph C Culberson \cite{PSpaceComplete} in 1997.
@@
 This shows PSPACE-completeness
+\textbf{Approaches:} There are loads of different researches regarding approaches to Sokoban.
+I decided which to show.
+Botea et al.
+found out that heuristics, while significantly improving chess AI, do not seem to help for planning the soko puzzle \cite{BoteaHeuristicsVsPlanning}.
+Planning would be a better approach.
+So to get there and create good plans like humans would, the solver will start meta-reasoning by recognizing rooms and tunnels.
+Then using those rooms and tunnels to create a plan where first a way is found to bring the crates from the room where theyre at to the target room.
+Then another plan is made to organize crates in a room.
+\newline
+\textbf{Macromoves:} Macromoves are a planning strategy that is so successful it is implemented in several good Sokoban solvers.
+The best example would be Rolling Stone, an excellent Sokoban solver, with a lot of optimizations implemented.
+One of the optimizations is the possibility to create macromoves for tunnels to reduce all states in the tunnel to one or two, reducing the search space severely.
+\newline
+\textbf{Deadlock Detection:} 
+%What is a Deadlock, with picture.
+Why do you detect those?
+How do you detect those?
+Better when you know the level or know basic deadlocks to look for.
+In the halls without tunnels it should be less hard
+\textbf{Relevance Cutting:}
+\newline
+\textbf{Lexicographical Ordering:}
+\newline
 

### 2018-08-20 BATexFiles/Draft.tex
Finished Section What is Sokoban and Section State of the Art - Some Pictures still missing (mockuppictures instead)

@@
 \section{Introduction}
-TODO: REWORK THIS
+%%%%TODO: REWORK THIS
 What do i do?
@@
 \subsection{Rules}
-\begin{center}
+\begin{figure}[h]
+\centering
 \includegraphics[scale=0.2]{SokobanLevel1}
-\newline
-\textbf{\textit{Fig.1: First of the Classical Sokoban Levels}}
-\end{center}
-Figure 1 illustrates the first Sokoban level of the original set of levels.
+\caption{First of the Classical Sokoban Levels}
+\label{fig:1}
+\end{figure}
+Figure \ref{fig:1} illustrates the first Sokoban level of the original set of levels.
 The game is played on a board that can be modelled as a two-dimensional grid.
@@
 \section{State of the Art}
+\paragraph*{•}
 \textbf{Complexity:} Gordon Wilfong published in 1991 that all motion planning problems with moving obstacles are NP-hard \cite{WilfongNPhard}.
@@
 %This shows PSPACE-hardness and Sokoban is known to be in PSPACE, so Sokoban is PSPACE-complete.
-\newline%He didnt answer the question whether Sokoban is also PSPACE-complete.
+%He didnt answer the question whether Sokoban is also PSPACE-complete.
 That was answered by Joseph C Culberson \cite{PSpaceComplete} in 1997.
@@
 This shows PSPACE-completeness
+\paragraph*{•}
 \textbf{Approaches:} There are loads of different researches regarding approaches to Sokoban.
@@
 Then another plan is made to organize crates in a room.
-\newline
+\paragraph*{•}
 \textbf{Macromoves:} Macromoves are a planning strategy that is so successful it is implemented in several good Sokoban solvers.
@@
 One of the optimizations is the possibility to create macromoves for tunnels to reduce all states in the tunnel to one or two, reducing the search space severely.
-\newline
-\textbf{Deadlock Detection:} 
+\paragraph*{•}
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.2]{flipped,deadlockexample}
+\caption{Deadlock example}
+\label{fig:2}
+\end{figure}
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.2]{deadlocktypes}
+\caption{Deadlock types}
+\label{fig:3}
+\end{figure}
+\textbf{Deadlock Detection:} Deadlocks are states from which no solution can be found.
+A trivial example for a deadlock (see Figure \ref{fig:2}) is a box pushed into a corner with no target under it.
+The box cannot be pulled, so it cannot be recovered from the corner, but to find a solution all boxes have to be covering a target.
+Detecting deadlocks the moment they appear is a very powerful optimization, for the reason that huge parts of the search tree can be eliminated early on, without further looking into them.
+The problem with this strategy is that deadlocks can be very hard to detect, e.g.
+when they consist of multiple boxes blocking each other in some kind of tunnelsystem.
+Deadlocks are easier to detect and remove from the search tree when the domain is restricted.
+When looking at levels with a rectangular shape there are four common deadlocks, cp.
+Fig.
+\ref{fig:3}.
+\paragraph*{•}
+%% A: Box in Corner, B: Doublebox on wall, C: Three boxes on corner, D: Four boxes building a square
 %What is a Deadlock, with picture.
@@
 In the halls without tunnels it should be less hard
-\textbf{Relevance Cutting:}
-\newline
-\textbf{Lexicographical Ordering:}
-\newline
-
-Heuristics doesnt seem to do a great deal for sokoban \cite{BoteaHeuristicsVsPlanning}.
+\textbf{Relevance Cutting:} A basic AI to solve a planning problem like Sokoban will look at all possible ways to move around the level, but some of those are dubious but legal or just plain irrelevant.
+In a lot of cases those dubious ways do not lead to an optimal solution, but of course a problem can be constructed where the optimal solution can only be found when looking at those dubious cases.
+So Junghanns and Schaeffer tried to teach their AI meta-reasoning to give it the ability to know when to look at those dubious ways and when to ignore them.
+This they did by creating an influence measurement and only taking moves that are influenced by prior moves, to prevent actions that are disembodied from all prior moves (and seem irrelevant).
+%Relevance Cutting tries to remove the problem of really stupid AI moves, like lrlrlrlrlrlrlr, by trying to teach the AI to know relevance, by teaching it influence of one move over another.
+This means that the moves can't be without any connection to each other and that the AI can realize that two parts of the puzzle have no influence on each other implying that order between the parts doesnt matter.
+the approach is statistic not dynamic which makes it rather simple and improvable.
+%Heuristics doesnt seem to do a great deal for sokoban \cite{BoteaHeuristicsVsPlanning}.
 It helped a lot with chess e.g., but doesnt cut it for sokoban.
@@
 Similarly you can also use relevance cutting.
-This means creating a measurement for relevance in between moves and picking moves that are relevant to prior moves (few exceptions in special cases).
-Backwards Approach and LexOrdering.
-Conclusion
+This means creating a measurement for relevance in between moves and picking moves that are relevant to prior moves (few exceptions in special cases).Backwards Approach and LexOrdering.Conclusion
 %Sokoban was researched to be PSPACE-complete and NP-hard.
@@
 %All in all a Plethora of different approaches to crack the problems of sokoban.
-
 %- mention other works on Sokoban
@@
 \section{My Work}
+In 
 what did i do?

### 2018-08-20 BATexFiles/Draft.tex
a little beauty correction for the tex file, had a lot of // for new lines, now replaced by \paragraph

@@
 \tableofcontents
+\newpage
 \section*{Abstract}
@@
 and finally evaluation and Discussion of Results.
+\newpage
 \section{What is Sokoban}
@@
 A level is usually satisfied when each crate is pushed on a target; also we can conclude that each target would be covered with a crate, because the amount of targets and crates is equal.
+\newpage
 \section{State of the Art}
@@
 %- whats the connection?
+\newpage
 \section{My Work}
+\paragraph*{•}
 In 
@@
 \subsubsection{Fact Format}
+\paragraph*{•}
 \includegraphics{Visualizer}
@@
 \subsubsection{Plain Encoding}
+\paragraph*{•}
 \lstinputlisting{../Encodings/SokobanPlainEncoding.lp}
@@
 \subsubsection{Incremental Encoding}
+\paragraph*{•}
 \lstinputlisting{../Encodings/SokobanPlainEncodingInc.lp}
@@
 \subsubsection{Advanced Encoding}
+\paragraph*{•}
 \lstinputlisting{../Encodings/SokobanPlainEncodingOpti.lp}
 \subsection{Evaluation}
+\paragraph*{•}
 \subsection{Experiments and Results}
+\paragraph*{•}
+\newpage
 \section{Discussion of Results}
+\paragraph*{•}
 compare chapter State of the Art with chapter My Work \cite{LifschitzASP}
@@
 \bibitem{LifschitzASP}
-Vladimir Lifschitz
-\textit{Answer Set Planning}
+Vladimir Lifschitz,
+\textit{Answer Set Planning},
 ICLP, 1999.
@@
 \bibitem{Lifschitz2002AnswerSP}
-Vladimir Lifschitz
-\textit{Answer set programming and plan generation}
+Vladimir Lifschitz,
+\textit{Answer set programming and plan generation},
 Artificial Intelligence, 2002, Vol. 138, p.39-54.
@@
 \bibitem{LifschitzWhatisASP}
-Vladimir Lifschitz
-\textit{What is Answer Set Programming?}
+Vladimir Lifschitz,
+\textit{What is Answer Set Programming?},
 AAAI, 2008.
@@
 \bibitem{Gebser2013RicochetRA}
-Martin Gebser, Holger Jost, Roland Kaminski, Philipp Obermeier, Orkunt Sabuncu, Torsten Schaub, Marius Thomas Lindauer
-\textit{Ricochet Robots: A Transverse ASP Benchmark}
+Martin Gebser, Holger Jost, Roland Kaminski, Philipp Obermeier, Orkunt Sabuncu, Torsten Schaub, Marius Thomas Lindauer,
+\textit{Ricochet Robots: A Transverse ASP Benchmark},
 LPMNR, 2013.
@@
 \bibitem{BoteaHeuristicsVsPlanning}
-Adi Botea, Martin Müller, Jonathan Schaeffer
-\textit{Using Abstraction for Planning in Sokoban}
+Adi Botea, Martin Müller, Jonathan Schaeffer,
+\textit{Using Abstraction for Planning in Sokoban},
 Computers and Games, 2002.
@@
 \bibitem{Frisch2006PropagationAF}
-Alan M. Frisch, Brahim Hnich, Zeynep Kiziltan, Ian Miguel, Toby Walsh
-\textit{Propagation algorithms for lexicographic ordering constraints}
+Alan M. Frisch, Brahim Hnich, Zeynep Kiziltan, Ian Miguel, Toby Walsh,
+\textit{Propagation algorithms for lexicographic ordering constraints},
 Artificial Intelligence, 2006, Vol. 170, p.803-834.
@@
 \bibitem{Dor1999SOKOBANAO}
-Dorit Dor, Uri Zwick
-\textit{SOKOBAN and other motion planning problems}
+Dorit Dor, Uri Zwick,
+\textit{SOKOBAN and other motion planning problems},
 Computational Geometry, 1999, Vol. 13, p.215-228.
@@
 \bibitem{Froleyks2017UsingAA}
-Nils Christian Froleyks and Tom{\'a}{\^s} Balyo
-\textit{Using an Algorithm Portfolio to Solve Sokoban}
+Nils Christian Froleyks and Tom{\'a}{\^s} Balyo,
+\textit{Using an Algorithm Portfolio to Solve Sokoban},
 SOCS, 2017.
@@
 \bibitem{PSpaceComplete}
-Joseph C. Culberson
-\textit{Sokoban is PSPACE-complete}
+Joseph C. Culberson,
+\textit{Sokoban is PSPACE-complete},
 1997.
@@
 \bibitem{SokoRelevanceCuts}
-Andreas Junghanns, Jonathan Schaeffer
-\textit{Sokoban: Improving the Search with Relevance Cuts}
+Andreas Junghanns, Jonathan Schaeffer,
+\textit{Sokoban: Improving the Search with Relevance Cuts},
 Journal of Theoretical Computing Science, 1999, Vol. 252, p.1-2.
@@
 \bibitem{SymmetryBreaking}
-Brahim Hnich, Zeynep Kiziltan, Toby Walsh
-\textit{Combining Symmetry Breaking with Other Constraints: Lexicographic Ordering with Sums}
+Brahim Hnich, Zeynep Kiziltan, Toby Walsh,
+\textit{Combining Symmetry Breaking with Other Constraints: Lexicographic Ordering with Sums},
 ISAIM, 2004.
@@
 \bibitem{WilfongNPhard}
-Gordon Wilfong
-\textit{Motion planning in the presence of movable obstacles}
+Gordon Wilfong,
+\textit{Motion planning in the presence of movable obstacles},
 Annals of Mathematics and Artificial Intelligence, 1991.

### 2018-08-26 BATexFiles/Draft.tex
no idea what changed but i'm trying a new approach in test2

@@
-\documentclass{article}
+\documentclass{report}
 \title{Efficient Planning for Sokoban with Answer Set Programming - A Case Study}
@@
 \usepackage{graphicx}
+\usepackage{titlesec}
+\titleformat{\chapter}
+  {\normalfont\LARGE\bfseries}{\thechapter}{1em}{}
 \graphicspath{{./images/}}
@@
 \tableofcontents
-\newpage
 \section*{Abstract}
 Sokoban approach with ASP planning $\rightarrow$ PlainEnc, IncEnc, Optimization (?)
-\section{Introduction}
-%%%%TODO: REWORK THIS
+\chapter{Introduction}
+%% Todo
 What do i do?
@@
 and finally evaluation and Discussion of Results.
-\newpage
-\section{What is Sokoban}
-\subsection{Background}
+\chapter{What is Sokoban}
+\section{Background}
 %% ZU KURZ, MIR FEHLT EIN BISSCHEN LORE (SOKOBAN HEISST WARENHAUSMANN, ENTWICKELT VON.., ...) UND EIN BISSCHEN FORMALITÄT (VGL \cite{Froleyks2017UsingAA})
@@
 %LOT OF RESEARCH REGARDING THAT (SO COMPLEX, SO EXEMPLARY, SO RELATABLE TO REAL LIFE AI PROBLEMS, SO EASY TO RESEARCH)
-\subsection{Rules}
+\section{Rules}
 \begin{figure}[h]
@@
 A level is usually satisfied when each crate is pushed on a target; also we can conclude that each target would be covered with a crate, because the amount of targets and crates is equal.
-\newpage
-\section{State of the Art}
+\chapter{State of the Art}
 \paragraph*{•}
@@
 %- whats the connection?
-\newpage
-\section{My Work}
+\chapter{My Work}
 \paragraph*{•}
-In 
-what did i do?
-what is my baseline?
-what changes from encoding to encoding
-which of the encodings works best in which kind of instance
-how do i evaluate that
-show my experiments and their results
-\subsection{Encodings}
-\subsubsection{Fact Format}
+In my research I focussed on simple approaches in very specific domains, to be able to compare the efficiency of different optimizations in those domains.
+The domains I researched are rectangular halls and slightly more crooked levels, for reference see Figure \ref{fig:domains}.
+Those maps had to be selfmade, so I had to find the horizon myself (by using incremental encodings).
+The basic optimizations are some approaches to deadlock detection, basic move-evaluation (don't take moves back), splitting x and y to reduce groundable variables, and i tried for lexordering the ways to reduce multiple grounded ways from A to B.
+%what did i do?what is my baseline?what changes from encoding to encoding?which of the encodings works best in which kind of instance?how do i evaluate that?show my experiments and their results!
+\section{Encodings}
+\subsection{Fact Format}
 \paragraph*{•}
+\begin{figure}[h]
+\centering
 \includegraphics{Visualizer}
+\caption{Instance 5x4 with 6 boxes}
+\label{fig:levelVis}
+\end{figure}
 \lstinputlisting{../InstancesClean/Inst-A-Z20-6Halle.lp}
-\subsubsection{Plain Encoding}
+In the code 
+\subsection{Plain Encoding}
 \paragraph*{•}
@@
 
-\subsubsection{Incremental Encoding}
+\subsection{Incremental Encoding}
 \paragraph*{•}
@@
 
-\subsubsection{Advanced Encoding}
+\subsection{Advanced Encoding}
 \paragraph*{•}
 \lstinputlisting{../Encodings/SokobanPlainEncodingOpti.lp}
-\subsection{Evaluation}
+\section{Evaluation}
 \paragraph*{•}
-\subsection{Experiments and Results}
+\section{Experiments and Results}
 \paragraph*{•}
-\newpage
-\section{Discussion of Results}
+\chapter{Discussion of Results}
 \paragraph*{•}

### 2018-09-10 BATexFiles/Draft.tex
Cleaning up the encodings and optimizations, not only in the folders but also internally

@@
 A level is usually satisfied when each crate is pushed on a target; also we can conclude that each target would be covered with a crate, because the amount of targets and crates is equal.
-\chapter{State of the Art}
+\chapter{Related Work}
 \paragraph*{•}
@@
 So Junghanns and Schaeffer tried to teach their AI meta-reasoning to give it the ability to know when to look at those dubious ways and when to ignore them.
-This they did by creating an influence measurement and only taking moves that are influenced by prior moves, to prevent actions that are disembodied from all prior moves (and seem irrelevant).
+This they did by creating an influence measurement and only taking moves that are influenced by prior moves, to prevent actions that are disembodied from all prior moves.
+This is used in the popular Sokoban Solver Rolling Stone which has an excellent record in approaching difficult Sokoban problems.
 %Relevance Cutting tries to remove the problem of really stupid AI moves, like lrlrlrlrlrlrlr, by trying to teach the AI to know relevance, by teaching it influence of one move over another.
@@
 \paragraph*{•}
-In my research I focussed on simple approaches in very specific domains, to be able to compare the efficiency of different optimizations in those domains.
+In my research I focused on simple approaches in very specific domains, to be able to compare the efficiency of different optimizations in those domains.
 The domains I researched are rectangular halls and slightly more crooked levels, for reference see Figure \ref{fig:domains}.
-Those maps had to be selfmade, so I had to find the horizon myself (by using incremental encodings).
+There are few common Sokoban domains that only look at levels with these qualities, so I designed a problemset myself.
+This meant that I had to find the horizon aswell (by using incremental encodings).
 The basic optimizations are some approaches to deadlock detection, basic move-evaluation (don't take moves back), splitting x and y to reduce groundable variables, and i tried for lexordering the ways to reduce multiple grounded ways from A to B.
@@
 \lstinputlisting{../InstancesClean/Inst-A-Z20-6Halle.lp}
-In the code 
+For illustration, consider the instance shown in Figure \ref{fig:levelVis} and the corresponding constants and facts in listing ABC.
+The turn limit is fixed by the constant \textit{horizon} and the level bounds are fixed by the constants \textit{dimX} and \textit{dimY}.
+The information about the level is stored in three different atoms:
+The \textit{init(field/2)} predicate stores the information about the levelstructure by holding the field coordinates of every field on which the player or a box can be placed.
+The \textit{init(at/3)} predicate holds the initial player- and box-positions.
+The first argument is the identifier of the object.
+A zero represents the player; all boxes have identifiers greater than zero.
+The last two arguments are the initial position of the object in the level.
+The \textit{init(target/2)} predicate has two arguments in levels containing no assignments.
+Here the two arguments represent the goal coordinates that have to be covered by a box to satisfy the level.
+In a level with assignments added there will be another argument in front of the other two referencing the identifier of the box that has to be placed on this goal position.
 \subsection{Plain Encoding}
 \paragraph*{•}
-\lstinputlisting{../Encodings/SokobanPlainEncoding.lp}
+\lstinputlisting{../EncodingsClean/SokobanPlainEncoding.lp}
 The first try to construct a naive encoding to solve sokoban-instances yielded a solution that works in a common way in ASP planning.
@@
 \paragraph*{•}
-\lstinputlisting{../Encodings/SokobanPlainEncodingInc.lp}
+\lstinputlisting{../EncodingsClean/SokobanIncEncoding.lp}
 The incremental encoding is similar to the plain encoding.
-The difference is the partition of the encoding into three parts: the \#program base in lines 3-5 which is independent of the step parameter t, the \#program step in lines 7-19 which is the cumulative part, collecting knowledge and the \#program check in lines 21-25 which is specific for each value of t (STEP ALSO IS THOUGH).
-This allows for gradually processing each time step and accumulating knowledge that persists even if the current step is not satisfiable.
+The difference is the partition of the encoding into three parts: the \#program base in lines 3-5 which is independent of the step parameter t, the \#program step in lines 7-19 which is the cumulative part, collecting knowledge and the \#program check in lines 21-25 which is loaded to check for each value of t and then unloaded for the next step.
+This allows for gradually processing each time step and accumulating knowledge that persists even if the current step is not satisfiable, greatly reducing the time in comparison to reloading the PlainEncoding everytime while incrementing the horizon by one.
 
@@
 \paragraph*{•}
-\lstinputlisting{../Encodings/SokobanPlainEncodingOpti.lp}
+\lstinputlisting{../EncodingsClean/Optimizations/Clutter.optimization.lp}
 \section{Evaluation}

### 2018-09-10 Stuff/Parser.py

@@ -1,5 +1,5 @@
 def read_map():
-    f = open('InstancesClean/Inst-A-Z20-6Halle.lp', 'r')
+    f = open('InstancesClean/Inst-A-Z20-4Halle.lp', 'r')
     sr = f.read()
     # f = open('atlist','r')
 

### 2018-09-10 Stuff/Visualizer.py

@@ -16,7 +16,7 @@ class Visualizer(QtWidgets.QApplication):
         my_grid = QtWidgets.QGridLayout()
         w.setLayout(my_grid)
 
-        f = open('InstancesClean/Inst-A-Z20-6Halle.lp', 'r')
+        f = open('InstancesClean/Inst-A-Z20-4Halle.lp', 'r')
         sr = f.read()
         n = get_max(sr)
         self.redraw(my_grid, w)

### 2018-09-12 BATexFiles/Draft.tex
added loads of text to the BA, tried around with encoding and goal conditions, made newinstance04 a little more complex

@@
 \section{Encodings}
+In the following, I will present the fact fromat and the first encodings I used to approach Sokoban solving in the input language of the ASP grounder gringo 3 \cite{Potassco}.
 \subsection{Fact Format}
@@
 \lstinputlisting{../InstancesClean/Inst-A-Z20-6Halle.lp}
-For illustration, consider the instance shown in Figure \ref{fig:levelVis} and the corresponding constants and facts in listing ABC.
-The turn limit is fixed by the constant \textit{horizon} and the level bounds are fixed by the constants \textit{dimX} and \textit{dimY}.
-The information about the level is stored in three different atoms:
-The \textit{init(field/2)} predicate stores the information about the levelstructure by holding the field coordinates of every field on which the player or a box can be placed.
-The \textit{init(at/3)} predicate holds the initial player- and box-positions.
-The first argument is the identifier of the object.
-A zero represents the player; all boxes have identifiers greater than zero.
-The last two arguments are the initial position of the object in the level.
-The \textit{init(target/2)} predicate has two arguments in levels containing no assignments.
-Here the two arguments represent the goal coordinates that have to be covered by a box to satisfy the level.
-In a level with assignments added there will be another argument in front of the other two referencing the identifier of the box that has to be placed on this goal position.
+\begin{lstlisting}[caption={Fact Format of my third level},label=lst:lvl1]
+\end{lstlisting}
+For illustration consider the level shown in Figure \ref{fig:levelVis}.
+The corresponding fact representation can be seen in Listing \ref{lst:lvl1}.
+The board size is fixed by the constants $dimX$ and $dimY$; the origin (1,1) is in the top left corner.
+The positions of player and boxes are stored in the at/3 predicate.
+The first argument represents the identifier.
+A zero stands for the player, while all boxes have numbers greater than zero.
+The second and third argument are the X and Y coordinates, signifying the initial position of the object.
+For instance init(at(0,1,2)) means the player is at the initial position (1,2).
+Because I am not exclusively looking at rectangular levels I decided to store all legal fields (meaning fields where a box or player can be placed) by using the predicate field/2.
+In rectangular levels the legal fields can be defined in one line.
+Here init(field(1..5,1..4)) (l.4) means the crossproduct of all the numbers in the pools (i.e.
+(1,1),..,(5,1),(1,2),..,(5,4)).
+The goal of the level is given in the target/2 predicate. Each coordinate given by the two arguments of a target/2 instance has to be covered by a box in the final gamestate to solve this level.
+Finally the constant horizon stores the information how many steps are needed to successfully satisfy the level. This horizon was determined by using the incremental encoding in section \ref{IncEnc}.
 \subsection{Plain Encoding}
 \paragraph*{•}
+\label{list:PlainEnc}
 \lstinputlisting{../EncodingsClean/SokobanPlainEncoding.lp}
-The first try to construct a naive encoding to solve sokoban-instances yielded a solution that works in a common way in ASP planning.
+\begin{lstlisting}[caption={Plain Encoding},label=list:PlainEnc]
+\end{lstlisting}
+The plain encoding seen in listing \ref{list:PlainEnc} uses a strategy common in ASP planning.
+A choice rule picks the moves (l.6) and the succeeding states are derived from those by using effect and frame axioms (ll.8-11).
+In the first few lines basic information is created, namely a $time/1$ predicate for each timestep (l.1), a $dir/2$ predicate for each of the four possible directions to move in (l.3) and an $at/4$ predicate to set the initial positions of objects as their position at time T=0.
+Additionally the encoding needs integrity constraints to ensure that the plans found are legal.
+There is one in line 13 that ensures that no object can move off the playable area and another in line 14 that prevents two object from being at the some position at the same time.
+In the lines 16 and 17 the goal condition is defined.
+The $box/2$ predicate is only created at the last timestep to specify each box position (X,Y).
+The constraint in line 17 stipulates that all targets have to have a corresponding box predicate with the same coordinates.
+The added directive \textbf{\#show at/4.} in line 21 allows for using clingos output as input to my Visualizer \textit{SokoViz}.
+This encoding allows to check for a plan of length equal to the horizon constant.
+This is of course not always optimal, e.g.
+when one wants to know a shortest plan to solve a level.
+To achieve an encoding that is able to, I will take advantage of incremental solving.
+$clingo$ provides an easy to use built-in incremental solving  and grounding mode.
+An encoding using the inc-mode can be seen in listing \ref{list:IncEnc}.
+It is very similar to the plain encoding from listing \ref{list:PlainEnc} in having the same structure of a choice rule picking the moves and effect and frame axioms finding the successive state.
+This encoding is split into three parts, which is critical for using the inc-mode.
+The \#program base part has only information that is created in the very beginning and will be unchanged throughout the whole grounding and solving process.
+This infomation is called \#static.
+The \#program step(t) part is the part accumulating knowledge, holding rules that will be collected and used for each step t and all following steps.
+The last part \#program check(t) holds volatile rules which will be checked once and discarded for the next timestep.
+Those rules are used to check for success and are no help to solving following steps.
+There is a minor change in the goal defining lines.
+In the plain encoding the horizon can be used to check efficiently only for the last timestep.
+In the incremental encoding I carry the time in the $box/3$ prediate to ensure that the constraint in line 21 only affects the target at time t.
+%The first try to construct a naive encoding to solve sokoban-instances yielded a solution that works in a common way in ASP planning.
+That is, guessing sequences of moves by a choice rule (line 7) and deriving the successive states by using effect and frame axioms (lines 9-13).
 In line 4 we initialize the information from the level instance with the time 0.
-In line 7 we state that there is one move at minimum and at maximum for each timestep.
+Line 9 calculates the new player-position, line 11 calculates the new position of the box in the players way and line 13 calculates the new position of boxes not touched by the player.
+Additionally we need some integrity constraints to ensure that no illegal solutions are found.
+The integrity constraint in line 15 ensures that at no time a box or player has a position outside of the playing field.
+The integrity constraint in line 17 ensures that at no time boxes or the player share one field.
+%In the lines 19-21 we define our win-condition to be that all targets in the instance have to be covered by boxes.
+%The Plain Encoding has the predicates, time/1, dir/2, at/4, move/3, uncovered/3, goal/3.
+%In line 7 we state that there is one move at minimum and at maximum for each timestep.
 This makes the ASP planner ground all sequences of moves, pick one for every timestep and infer every successive state from the moves it picks.
 In lines 9-13 we calculate the successor state from the move picked in line 7.
-Line 9 calculates the new player-position, line 11 calculates the new position of the box in the players way and line 13 calculates the new position of boxes not touched by the player.
-Additionally we need some integrity constraints to prevent illegal states.
-The integrity constraint in line 15 ensures that at no time a box or player has a position outside of the playing field.
-The integrity constraint in line 17 ensures that at no time boxes or the player share one field.
-In the lines 19-21 we define our win-condition to be that all targets in the instance have to be covered by boxes.
-The Plain Encoding has the predicates, time/1, dir/2, at/4, move/3, uncovered/3, goal/3
-The encoding works in a common way in ASP planning, by guessing the sequence of actions and inferring the successor states by effect axioms.
+%The encoding works in a common way in ASP planning, by guessing the sequence of actions and inferring the successor states by effect axioms.
 There is a time/1 predicate for every timestep and a at/4 predicate for every box and the player at every timestep.
 
-\subsection{Incremental Encoding}
+\subsection{Incremental Encoding} \label{IncEnc}
 \paragraph*{•}
 \lstinputlisting{../EncodingsClean/SokobanIncEncoding.lp}
-The incremental encoding is similar to the plain encoding.
+\begin{lstlisting}[caption={Incremental Encoding},label=list:IncEnc]
+\end{lstlisting}
+%The incremental encoding is similar to the plain encoding.
 The difference is the partition of the encoding into three parts: the \#program base in lines 3-5 which is independent of the step parameter t, the \#program step in lines 7-19 which is the cumulative part, collecting knowledge and the \#program check in lines 21-25 which is loaded to check for each value of t and then unloaded for the next step.
@@
 
+\bibitem{Potassco}
+Martin Gebser, Roland Kaminski, Benjamin Kaufmann, Marius Lindauer, Max Ostrowski, Javier Romero, Torsten Schaub, Sven Thiele,
+\textit{Potassco User Guide},
+Second Edition, 2015.
+
 \bibitem{WilfongNPhard}

### 2018-09-18 BATexFiles/Draft.tex
Added lots of text

@@
 \usepackage{titlesec}
-\titleformat{\chapter}
-  {\normalfont\LARGE\bfseries}{\thechapter}{1em}{}
+%\titleformat{\chapter}
+%  {\normalfont\LARGE\bfseries}{\thechapter}{1em}{}
 \graphicspath{{./images/}}
 \usepackage{color}
-\lstset{numbers=left,
-literate = {-}{-}1,
-basicstyle=\footnotesize,
-numbersep=5pt, backgroundcolor=\color{white},
-captionpos=b,keywordstyle=\color{blue},
-commentstyle=\color{green}}
+%\lstset{numbers=left,
+%literate = {-}{-}1,
+%basicstyle=\footnotesize,
+%numbersep=5pt, backgroundcolor=\color{white},
+%captionpos=b,keywordstyle=\color{blue},
+%commentstyle=\color{green}}
 \begin{document}
@@
 \section*{Abstract}
-Sokoban approach with ASP planning $\rightarrow$ PlainEnc, IncEnc, Optimization (?)
 \chapter{Introduction}
-%% Todo
-What do i do?
-I'm working on the Sokoban problem and trying to do my best to find a fast and easy solution to it.
-I'm not working on any level but just the ones meeting some criteria.
-I'm picking qualitative and quantitative characteristics and deciding on some values to examine.
-My encodings are supposed to work better on some of the instances and worse on others.
-What rules describe those fluctuations?
-What assumptions can be made?
-Is there a trend when comparing runtimes to my qualitative and quantitative characteristics?
-E.g.
-you would expect the runtime to go up when the instance's size grows or when the amount of boxes in the level grows.
-In this thesis I will work on optimizing ASP encodings for the game Sokoban towards different instances.
-First Ill briefly explain what Sokoban is and how it is played, then Ill give a short introduction on comparable/similar/other scientific studies concerning sokoban.
-Then Ill go over my work, ...
-and finally evaluation and Discussion of Results.
 \chapter{What is Sokoban}
 \section{Background}
-%% ZU KURZ, MIR FEHLT EIN BISSCHEN LORE (SOKOBAN HEISST WARENHAUSMANN, ENTWICKELT VON.., ...) UND EIN BISSCHEN FORMALITÄT (VGL \cite{Froleyks2017UsingAA})
 Sokoban is a logistic planning puzzle game from Japan.
@@
 This makes Sokoban a very interesting subject for research regarding planning problems.
-%LOT OF RESEARCH REGARDING THAT (SO COMPLEX, SO EXEMPLARY, SO RELATABLE TO REAL LIFE AI PROBLEMS, SO EASY TO RESEARCH)
 \section{Rules}
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 There are a lot of possibilities to create a deadlock (a gamestate from which no solution can be found), the branching factor is large and the optimal solution can be very long \cite{BoteaHeuristicsVsPlanning}.
-
-%Sokoban is a logistics game, where the player controls a character in a warehouse.
-The goal of the game is to move all crates in the warehouse on one of the target fields.
-The rules state that crates cannot be stacked on one field and crates cannot be pushed if there is a crate or a wall blocking the way.
-Each player move is a step into one of the four directions in a two-dimensional grid.
-If the player would collide with a crate, the crate is pushed into the direction of the player-movement.
-A level is usually satisfied when each crate is pushed on a target; also we can conclude that each target would be covered with a crate, because the amount of targets and crates is equal.
 \chapter{Related Work}
-\paragraph*{•}
-\textbf{Complexity:} Gordon Wilfong published in 1991 that all motion planning problems with moving obstacles are NP-hard \cite{WilfongNPhard}.
+\paragraph{Complexity:} Gordon Wilfong published in 1991 that all motion planning problems with moving obstacles are NP-hard \cite{WilfongNPhard}.
 He proved this by reducing $3$-$Sat$ to such a problem.
@@
 He found a transformation algorithm to transform an LBA into a level of Sokoban, with the solution to the level having a number of steps in $\Theta (|w|+t(|w|))$, where $w$ is the input and $t(|w|)$ is the amount of transitions the LBA made.
-%This shows PSPACE-hardness and Sokoban is known to be in PSPACE, so Sokoban is PSPACE-complete.
-%He didnt answer the question whether Sokoban is also PSPACE-complete.
-That was answered by Joseph C Culberson \cite{PSpaceComplete} in 1997.
-He showed that it is possible to emulate a linear bounded automaton by constructing a level.
-The solution to this constructed level, if there is any, will have a length in $\Theta (|w|+t(|w|))$, where $w$ is the input and $t(|w|)$ is the amount of transitions the LBA made.
-This shows PSPACE-completeness
-\paragraph*{•}
-\textbf{Approaches:} There are loads of different researches regarding approaches to Sokoban.
+\paragraph{Approaches:} There are loads of different researches regarding approaches to Sokoban.
 I decided which to show.
@@
 Then another plan is made to organize crates in a room.
-\paragraph*{•}
-\textbf{Macromoves:} Macromoves are a planning strategy that is so successful it is implemented in several good Sokoban solvers.
+\paragraph{Macromoves:} Macromoves are a planning strategy that is so successful it is implemented in several good Sokoban solvers.
 The best example would be Rolling Stone, an excellent Sokoban solver, with a lot of optimizations implemented.
 One of the optimizations is the possibility to create macromoves for tunnels to reduce all states in the tunnel to one or two, reducing the search space severely.
-\paragraph*{•}
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 \end{figure}
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 \end{figure}
-\textbf{Deadlock Detection:} Deadlocks are states from which no solution can be found.
+\paragraph{Deadlock Detection:} Deadlocks are states from which no solution can be found.
 A trivial example for a deadlock (see Figure \ref{fig:2}) is a box pushed into a corner with no target under it.
@@
 \ref{fig:3}.
-\paragraph*{•}
-%% A: Box in Corner, B: Doublebox on wall, C: Three boxes on corner, D: Four boxes building a square
-%What is a Deadlock, with picture.
-Why do you detect those?
-How do you detect those?
-Better when you know the level or know basic deadlocks to look for.
-In the halls without tunnels it should be less hard
-\textbf{Relevance Cutting:} A basic AI to solve a planning problem like Sokoban will look at all possible ways to move around the level, but some of those are dubious but legal or just plain irrelevant.
+\paragraph{Relevance Cutting:} A basic AI to solve a planning problem like Sokoban will look at all possible ways to move around the level, but some of those are dubious but legal or just plain irrelevant.
 In a lot of cases those dubious ways do not lead to an optimal solution, but of course a problem can be constructed where the optimal solution can only be found when looking at those dubious cases.
@@
 This they did by creating an influence measurement and only taking moves that are influenced by prior moves, to prevent actions that are disembodied from all prior moves.
-This is used in the popular Sokoban Solver Rolling Stone which has an excellent record in approaching difficult Sokoban problems.
-%Relevance Cutting tries to remove the problem of really stupid AI moves, like lrlrlrlrlrlrlr, by trying to teach the AI to know relevance, by teaching it influence of one move over another.
-This means that the moves can't be without any connection to each other and that the AI can realize that two parts of the puzzle have no influence on each other implying that order between the parts doesnt matter.
-the approach is statistic not dynamic which makes it rather simple and improvable.
-%Heuristics doesnt seem to do a great deal for sokoban \cite{BoteaHeuristicsVsPlanning}.
-It helped a lot with chess e.g., but doesnt cut it for sokoban.
-Approach the hard domain of sokoban in another way.
-Macromoves a way that really works well and is used in practice, also deadlock detection (WHAT IS A DEADLOCK - stone in corner and no goal -> no solutions) (obviously best if you know the specific level).
-Reference Rolling stone (source not available (?)).
-Furthermore Sokoban can be abstracted into rooms and tunnels, making it into subproblems which should be easier to solve, tunnels for example, are mostly immediately solved.
-Similarly you can also use relevance cutting.
-This means creating a measurement for relevance in between moves and picking moves that are relevant to prior moves (few exceptions in special cases).Backwards Approach and LexOrdering.Conclusion
-%Sokoban was researched to be PSPACE-complete and NP-hard.
-PSPACE-complete was done by constructing levels that are PSPACE-complete (look at the work again!!).
-Heuristic seems to be of little use, so go for planning instead [botea]
-%Sokoban can be abstracted into subproblems to get more smaller problems, greatly reducing searchspace.
-%This also makes sense in relevance cutting, which looks at relations between boxes to find groups that behave together and cleanly separate those.
-%Alternatively Backwards Approach
-%Or LexOrdering
-%All in all a Plethora of different approaches to crack the problems of sokoban.
-%- mention other works on Sokoban
-%- what tests?
-%- what results?
-%- whats the connection?
+This is used in the popular Sokoban solver Rolling Stone which has an excellent record in solving Sokoban problems.
 \chapter{My Work}
-\paragraph*{•}
-In my research I focused on simple approaches in very specific domains, to be able to compare the efficiency of different optimizations in those domains.
-The domains I researched are rectangular halls and slightly more crooked levels, for reference see Figure \ref{fig:domains}.
-There are few common Sokoban domains that only look at levels with these qualities, so I designed a problemset myself.
-This meant that I had to find the horizon aswell (by using incremental encodings).
-The basic optimizations are some approaches to deadlock detection, basic move-evaluation (don't take moves back), splitting x and y to reduce groundable variables, and i tried for lexordering the ways to reduce multiple grounded ways from A to B.
-%what did i do?what is my baseline?what changes from encoding to encoding?which of the encodings works best in which kind of instance?how do i evaluate that?show my experiments and their results!
+IM DOING ASP AND THAT IS DIFFERENT TO MOST OTHER APPROACHES I LOOKED AT
+\cite{BoteaHeuristicsVsPlanning} \cite{Dor1999SOKOBANAO} \cite{Froleyks2017UsingAA} \cite{SokoRelevanceCuts}
 \section{Encodings}
-In the following, I will present the fact fromat and the first encodings I used to approach Sokoban solving in the input language of the ASP grounder gringo 3 \cite{Potassco}.
+In the following, I will present the fact format and the first encodings I used to approach Sokoban solving in the input language of the ASP grounder gringo 3 \cite{Potassco}.
 \subsection{Fact Format}
-\paragraph*{•}
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 \subsection{Plain Encoding}
-\paragraph*{•}
-\label{list:PlainEnc}
 \lstinputlisting{../EncodingsClean/SokobanPlainEncoding.lp}
@@
 To achieve an encoding that is able to, I will take advantage of incremental solving.
+\subsection{Incremental Encoding} \label{IncEnc}
+\lstinputlisting{../EncodingsClean/SokobanIncEncoding.lp}
+\begin{lstlisting}[caption={Incremental Encoding},label=list:IncEnc]
+\end{lstlisting}
 $clingo$ provides an easy to use built-in incremental solving  and grounding mode.
@@
 In the incremental encoding I carry the time in the $box/3$ prediate to ensure that the constraint in line 21 only affects the target at time t.
-%The first try to construct a naive encoding to solve sokoban-instances yielded a solution that works in a common way in ASP planning.
-That is, guessing sequences of moves by a choice rule (line 7) and deriving the successive states by using effect and frame axioms (lines 9-13).
-In line 4 we initialize the information from the level instance with the time 0.
-Line 9 calculates the new player-position, line 11 calculates the new position of the box in the players way and line 13 calculates the new position of boxes not touched by the player.
-Additionally we need some integrity constraints to ensure that no illegal solutions are found.
-The integrity constraint in line 15 ensures that at no time a box or player has a position outside of the playing field.
-The integrity constraint in line 17 ensures that at no time boxes or the player share one field.
-%In the lines 19-21 we define our win-condition to be that all targets in the instance have to be covered by boxes.
-%The Plain Encoding has the predicates, time/1, dir/2, at/4, move/3, uncovered/3, goal/3.
-%In line 7 we state that there is one move at minimum and at maximum for each timestep.
-This makes the ASP planner ground all sequences of moves, pick one for every timestep and infer every successive state from the moves it picks.
-In lines 9-13 we calculate the successor state from the move picked in line 7.
-%The encoding works in a common way in ASP planning, by guessing the sequence of actions and inferring the successor states by effect axioms.
-There is a time/1 predicate for every timestep and a at/4 predicate for every box and the player at every timestep.
 
-\subsection{Incremental Encoding} \label{IncEnc}
-\paragraph*{•}
-\lstinputlisting{../EncodingsClean/SokobanIncEncoding.lp}
-\begin{lstlisting}[caption={Incremental Encoding},label=list:IncEnc]
+\subsection{Split Encoding}
+\lstinputlisting{../EncodingsClean/SokobanSplitEncoding.lp}
+\begin{lstlisting}[caption={Split Encoding},label=list:SplitEnc]
 \end{lstlisting}
-%The incremental encoding is similar to the plain encoding.
-The difference is the partition of the encoding into three parts: the \#program base in lines 3-5 which is independent of the step parameter t, the \#program step in lines 7-19 which is the cumulative part, collecting knowledge and the \#program check in lines 21-25 which is loaded to check for each value of t and then unloaded for the next step.
-This allows for gradually processing each time step and accumulating knowledge that persists even if the current step is not satisfiable, greatly reducing the time in comparison to reloading the PlainEncoding everytime while incrementing the horizon by one.
+The split encoding uses a technique common in ASP planning.
+When all coordinates are split into a vertical and a horizontal information the grounder does not have to ground all combinations of coordinates ($X * Y$) but only the sum of coordinates ($X + Y$).
+This has the advantage of less ground instances, but the disadvantage of a harder solving process.
+Just compare the ramifications in the plain encoding those in the split encoding.
+The plain encoding can deal with all ramifications in just three rules, while the split encoding needs ten rules.
+Obviously when splitting the coordinates into horizontal and vertical the rules deriving the new coordinates have to be twice as many.
+The other problem is that it is harder to find out whether two objects meet each other at a time.
+That is why the frame axiom, signifying that boxes that are not touched do not move, has six lines, where it had only one line before.
+This is why this encoding has a performance problem.
+The last few changes are a second line for the initial readout of the level in lines 4 and 5, a few updates to the integrity constraints in lines 24 and 25 and a slightly different goal definition using the new coordinates.
+%Here is another encoding, following a common ASP planning technique that aims to reduce ground instances.
+By splitting the horizontal and vertical coordinates, the program does not have to ground all combinations ($dimX * dimY$) of coordinates, but just each coordinate by itself (totaling $dimX + dimY$).% only grounds $(dimX + dimY) * horizon * N$ different positions instead of $dimX * dimY * horizon * N$.
 
-\subsection{Advanced Encoding}
-\paragraph*{•}
+\subsection{Optimizations}
+
+\subsubsection{Deadlock Detection}
+
+\lstinputlisting{../EncodingsClean/Optimizations/ForbiddenFields.optimization.lp}
+\begin{lstlisting}[caption={Forbidden Fields},label=list:Forbidden]
+\end{lstlisting}
+
+Whenever an AI tries to solve a planning problem with a naive encoding, it always has to guess all possible combinations of moves and just has to play them through.
+There are a lot of moves that instantly destroy any chanco in ever satisfiying the level afterwards.
+Those gamestates from which no solution can be found are called deadlocks.
+A search tree can be greatly reduced if one were to remove all deadlocked branches.
+The difficulty is that some deadlocks are very hard to notice and sometimes it takes more time elimintaing the deadlock than to explore the search tree without eliminating.
+This is why the deadlock detection approach takes only obvious deadlocks.
+There are three different approaches I tried to remove deadlocks.
+The Forbidden Fields optimization (listing \ref{list:Forbidden}) works by generating a list of forbidden fields.
+Those fields are creating deadlocks when the player pushes a box onto them.
+This approach has the benefit of having a lot of that one can exclude from all solution pathways.
+Those fields are deduced in the \#static part of the program, i.e.
+they are only grounded once and do not change over the timesteps in a level instance.
+The disadvantage of this approach is the rather long time it takes to find all forbidden fields.
+Furthermore the approach works best in levels that do not have any restrictions on how they are constructed.
+As I am looking into levels having a very specific structure, see \ref{chapter 1 or 2 introduction}, this implementation is not the best approach.
+In rectangular levels there is an easier way to express a very similar deadlock detection, see next paragraph.
+
+\lstinputlisting{../EncodingsClean/Optimizations/WallBoxTargetCheck.optimization.lp}
+\begin{lstlisting}[caption={Deadlock check},label=list:WBT]
+\end{lstlisting}
+
+This approach can be seen in listing \ref{list:WBT}.
+This optimization works best in rectangular levels by simplifying the forbidden check for this special case.
+The implementaion does not look at tunnels or similar structures but only checks for the outermost lines and rows.
+When a box is pushed into one of those lines or rows one will not be able to remove it from that line or row.
+So the idea is to compare the number of boxes in an outermost line or row to the number of targets in it.
+If the number of boxes exceeds the number of targets the level cannot be satisfied anymore.
+The result is a very slim check which can help to greatly reduce a search tree.
+
 \lstinputlisting{../EncodingsClean/Optimizations/Clutter.optimization.lp}
+\begin{lstlisting}[caption={Clutter},label=list:Clutter]
+\end{lstlisting}
+
+My third approach to deadlock detection (see listing \ref{list:Clutter}) recognizes a "clutter" of boxes (see Fig.
+\ref{fig:clutter}).
+This means a 2x2 square where each field is either a box or a wall, which means that all those boxes would have to stay in their places.
+In line 8 of the optimization we can see that the box D does not have a target underneath which qualifies the situation to be a deadlock.
+This type of deadlock happens a lot in crammed levels with a high box to field ratio.
+Those very common deadlocks are clearly most important to find and remove to reduce the search tree efficiently.
+
+\subsubsection{Move Evaluation}
+
+In the following I will show two more optimizations working with move evaluation. This means eliminating movies that will not be relevant for a shortest plan.
+
+\lstinputlisting{../EncodingsClean/Optimizations/MovesOutOfBounds.optimization.lp}
+\begin{lstlisting}[caption={Moves out of Bounds},label=list:MOOB]
+\end{lstlisting}
+
+The optimization in listing \ref{list:MOOB} is very simple but can remove a lot of moves from grounding, which would be removed after ramifications when checking integrity constraints otherwise.
+This is more important in levels with a very long shortest plan.
+The optimization works by preventing all moves that would lead the player of the playable area, for instance a move to the left while the player is at the leftmost position of the level.
+
+\lstinputlisting{../EncodingsClean/Optimizations/NoTakebacks.optimization.lp}
+\begin{lstlisting}[caption={No Taking back Moves},label=list:NoTake]
+\end{lstlisting}
+
+The last move evaluation seen in listing \ref{list:NoTake} tries to recognize moves that have no effect.
+This is done by checking for move instances $move(DX,DY,T)$ and $move(-DX,-DY,T-1)$ where no box was moved in timestep T-1.
+This would imply that the player moved back and forth without actually changing the gamestate, which means the plan is not relevant for finding the shortest plan in the level instance.
+
 \section{Evaluation}
-\paragraph*{•}
+
 \section{Experiments and Results}
-\paragraph*{•}
+
 \chapter{Discussion of Results}
-\paragraph*{•}
 compare chapter State of the Art with chapter My Work \cite{LifschitzASP}

### 2018-09-18 BATexFiles/Draft.tex
Textversion 1 abgeschlossen

@@
 \usepackage{titlesec}
-%\titleformat{\chapter}
-%  {\normalfont\LARGE\bfseries}{\thechapter}{1em}{}
 \graphicspath{{./images/}}
@@
 This makes Sokoban a very interesting subject for research regarding planning problems.
-\section{Rules}
+\section{Rules} \label{IntroStr}
 \begin{figure}[ht]
@@
 \chapter{My Work}
-IM DOING ASP AND THAT IS DIFFERENT TO MOST OTHER APPROACHES I LOOKED AT
+In my work I decided to research Sokoban planning with Answer Set Programming (ASP), a logic programming system designed to return answer sets to queries \cite{LifschitzASP}.
+I created different basic encodings and optimizations in a modular fashion.
+Those I will present to you in this chapter.
+Additionally I created a level portfolio, corresponding to different qualitative properties, e.g.
+level dimensions or number of boxes, and quantitative properties e.g.
+a seperation of the level into a zone where all the boxes are placed and a zone where all the targets are placed.
+The levels have two different types of structures.
+The first type is a rectangular level with the X and Y dimension differing by 1.
+The second type is created by connecting 2x2 fields at their edges, see \ref{fig:levelCrooked}.
+%. But instead of Prolog ASP is capable of producing answer sets to a query
 \cite{BoteaHeuristicsVsPlanning} \cite{Dor1999SOKOBANAO} \cite{Froleyks2017UsingAA} \cite{SokoRelevanceCuts}
@@
 The corresponding fact representation can be seen in Listing \ref{lst:lvl1}.
-The board size is fixed by the constants $dimX$ and $dimY$; the origin (1,1) is in the top left corner.
+The board size is fixed by the constants $dimX$ and $dimY$ and the origin (1,1) is in the top left corner.
 The positions of player and boxes are stored in the at/3 predicate.
-The first argument represents the identifier.
+Here the first argument represents the identifier.
 A zero stands for the player, while all boxes have numbers greater than zero.
@@
 The other problem is that it is harder to find out whether two objects meet each other at a time.
-That is why the frame axiom, signifying that boxes that are not touched do not move, has six lines, where it had only one line before.
-This is why this encoding has a performance problem.
-The last few changes are a second line for the initial readout of the level in lines 4 and 5, a few updates to the integrity constraints in lines 24 and 25 and a slightly different goal definition using the new coordinates.
+That is why the two frame axioms, signifying that boxes that are not touched do not move, have aggregates (ll.17-20).
+Those are used for simplifying complex rules; the first rule without the aggregates can be seen in listing \ref{list:BadFrame}.
+\lstinputlisting{../EncodingsClean/BadSplitFrameAxiom.lp}
+\begin{lstlisting}[caption={Frame Axiom X},label=list:BadFrame]
+\end{lstlisting}
+The last few changes are a second line for the initial readout of the level in lines 4 and 5, a few updates to the integrity constraints in lines 24 and 25 and a slightly different goal definition using the split coordinates.
 %Here is another encoding, following a common ASP planning technique that aims to reduce ground instances.
@@
 
-Whenever an AI tries to solve a planning problem with a naive encoding, it always has to guess all possible combinations of moves and just has to play them through.
-There are a lot of moves that instantly destroy any chanco in ever satisfiying the level afterwards.
-Those gamestates from which no solution can be found are called deadlocks.
+Whenever an AI tries to solve a planning problem with a naive encoding, it does not have the meta reasoning a human would have \cite{BoteaHeuristicsVsPlanning}, \cite{SokoRelevanceCuts}.
+This means it cannot see whether a move is worthless before it calculates its effect.
+So it has to play through all possible sequences of moves, even those where the first few moves already destroy any chance to satisfy the level.
+Those gamestates from which no solution can be reached are called deadlocks.
 A search tree can be greatly reduced if one were to remove all deadlocked branches.
-The difficulty is that some deadlocks are very hard to notice and sometimes it takes more time elimintaing the deadlock than to explore the search tree without eliminating.
-This is why the deadlock detection approach takes only obvious deadlocks.
+The difficulty is that some deadlocks are very hard to notice and sometimes it takes more time eliminating the deadlock than to explore the search tree without eliminating.
+This is why my deadlock detection approach removes only deadlocks, that are easy to check for.
 There are three different approaches I tried to remove deadlocks.
@@
 Those fields are creating deadlocks when the player pushes a box onto them.
-This approach has the benefit of having a lot of that one can exclude from all solution pathways.
+This approach has the benefit of having a lot of fields that one can exclude from all solution pathways.
 Those fields are deduced in the \#static part of the program, i.e.
@@
 The disadvantage of this approach is the rather long time it takes to find all forbidden fields.
-Furthermore the approach works best in levels that do not have any restrictions on how they are constructed.
-As I am looking into levels having a very specific structure, see \ref{chapter 1 or 2 introduction}, this implementation is not the best approach.
+Furthermore the approach is suited for levels that do not have any restrictions on how they are constructed.
+As I am looking into levels having a very specific structure (see \ref{IntroStr}), this implementation is not the best approach.
 In rectangular levels there is an easier way to express a very similar deadlock detection, see next paragraph.
@@
 \ref{fig:clutter}).
-This means a 2x2 square where each field is either a box or a wall, which means that all those boxes would have to stay in their places.
-In line 8 of the optimization we can see that the box D does not have a target underneath which qualifies the situation to be a deadlock.
+This is a 2x2 square where each field is either a box or a wall, which would imply that all those boxes would have to stay in their places.
+In line 8 of the optimization we can see that the box D does not have a target underneath, which qualifies the situation to be a deadlock.
 This type of deadlock happens a lot in crammed levels with a high box to field ratio.
-Those very common deadlocks are clearly most important to find and remove to reduce the search tree efficiently.
+Those very common deadlocks are very important to find and remove to reduce the search tree in those levels efficiently.
 
@@
 The optimization in listing \ref{list:MOOB} is very simple but can remove a lot of moves from grounding, which would be removed after ramifications when checking integrity constraints otherwise.
-This is more important in levels with a very long shortest plan.
-The optimization works by preventing all moves that would lead the player of the playable area, for instance a move to the left while the player is at the leftmost position of the level.
+This is more important in levels with a high horizon constant, i.e.
+a long shortest plan.
+The optimization works by preventing all moves that would lead the player of the playable area, for instance a move to the left, while the player is at the leftmost position of the level.
 

### 2018-09-25 bashscript.sh
changed a lot, trying to get this done

@@ -1,0 +1,25 @@
+#!/bin/bash
+clingo 
+for enc in "EncodingsClean/Test"/*.lp
+do
+	for inst in "InstancesClean/Test"/*.lp
+	do
+		clingo $enc $inst --stats --time-limit=1800 >> results.txt
+	done
+done
+
+for entry in "EncodingsClean/Assignments"/*.lp
+do
+	echo "$entry"
+done
+
+for entry in "EncodingsClean/Optimizations"/*.lp
+do
+	echo "$entry"
+done
+
+
+for entry in "InstancesClean/Assignments"/*.lp
+do
+	echo "$entry"
+done

### 2018-09-25 bashscript.sh
going into testing

@@ -1,25 +1,58 @@
 #!/bin/bash
-clingo 
-for enc in "EncodingsClean/Test"/*.lp
+plain="EncodingsClean/SokobanPlainEncoding.lp"
+inc="EncodingsClean/SokobanIncEncoding.lp"
+split="EncodingsClean/SokobanSplitEncoding.lp"
+splitinc="EncodingsClean/SokobanSplitIncEncoding.lp"
+plainass="EncodingsClean/Assignments/SokobanPlainEncodingAssignments.lp"
+incass="EncodingsClean/Assignments/SokobanIncEncodingAssignments.lp"
+splitass="EncodingsClean/Assignments/SokobanSplitEncodingAssignments.lp"
+splitincass="EncodingsClean/Assignments/SokobanSplitIncEncodingAssignments.lp"
+
+clutter="EncodingsClean/Optimizations/Clutter.optimization.lp"
+forfie="EncodingsClean/Optimizations/ForbiddenFields.optimization.lp"
+moob="EncodingsClean/Optimizations/MovesOutOfBounds.optimization.lp"
+ntb="EncodingsClean/Optimizations/NoTakebacks.optimization.lp"
+wbtc="EncodingsClean/Optimizations/WallBoxTargetCheck.optimization.lp"
+
+spclutter="EncodingsClean/Optimizations Split/SplitClutter.optimization.lp"
+spforfie="EncodingsClean/Optimizations Split/SplitForbiddenFields.optimization.lp"
+spmoob="EncodingsClean/Optimizations Split/SplitMovesOutOfBounds.optimization.lp"
+spntb="EncodingsClean/Optimizations Split/SplitNoTakebacks.optimization.lp"
+spwbtc="EncodingsClean/Optimizations Split/SplitWallBoxTargetCheck.optimization.lp"
+
+index=0
+
+for inst in "InstancesClean"/Inst-A-Z*.lp
 do
-	for inst in "InstancesClean/Test"/*.lp
-	do
-		clingo $enc $inst --stats --time-limit=1800 >> results.txt
-	done
+	index=$(($index+1))
+	clingo $plain $inst $clutter --stats -q2 --time-limit=900 >> resultsC${index}.txt
+	clingo $plain $inst $forfie --stats -q2 --time-limit=900 >> resultsFF${index}.txt
+	clingo $plain $inst $moob --stats -q2 --time-limit=900 >> resultsMOOB${index}.txt
[…]

### 2018-10-02 BATexFiles/Draft.tex
changed a lot, trying to get this done
asd

@@
 \paragraph{Macromoves:} Macromoves are a planning strategy that is so successful it is implemented in several good Sokoban solvers.
-The best example would be Rolling Stone, an excellent Sokoban solver, with a lot of optimizations implemented.
-One of the optimizations is the possibility to create macromoves for tunnels to reduce all states in the tunnel to one or two, reducing the search space severely.
+The best example would be Rolling Stone, an excellent Sokoban solver, with a lot of optimizations and strategies implemented.
+One of the strategies is the possibility to create macromoves for tunnels to collapse all states in the tunnel into one or two different states, reducing the search space severely \cite{SokoRelevanceCuts}.
 \begin{figure}[ht]
@@
 The box cannot be pulled, so it cannot be recovered from the corner, but to find a solution all boxes have to be covering a target.
-Detecting deadlocks the moment they appear is a very powerful optimization, for the reason that huge parts of the search tree can be eliminated early on, without further looking into them.
+Detecting deadlocks the moment they appear is a very powerful solution strategy, for the reason that huge parts of the search tree can be eliminated early on, without further looking into them \cite{SokoRelevanceCuts}.
 The problem with this strategy is that deadlocks can be very hard to detect, e.g.
@@
 Fig.
-\ref{fig:3}.
+\ref{fig:3} .
 \paragraph{Relevance Cutting:} A basic AI to solve a planning problem like Sokoban will look at all possible ways to move around the level, but some of those are dubious but legal or just plain irrelevant.
@@
 This they did by creating an influence measurement and only taking moves that are influenced by prior moves, to prevent actions that are disembodied from all prior moves.
-This is used in the popular Sokoban solver Rolling Stone which has an excellent record in solving Sokoban problems.
+This is used in the popular Sokoban solver Rolling Stone which has an excellent record in solving Sokoban problems \cite{SokoRelevanceCuts}.
 \chapter{My Work}
 In my work I decided to research Sokoban planning with Answer Set Programming (ASP), a logic programming system designed to return answer sets to queries \cite{LifschitzASP}.
-I created different basic encodings and optimizations in a modular fashion.
+I created different basic encodings and strategy implementations in a modular fashion.
 Those I will present to you in this chapter.
@@
 The second type is created by connecting 2x2 fields at their edges, see \ref{fig:levelCrooked}.
-%. But instead of Prolog ASP is capable of producing answer sets to a query
-\cite{BoteaHeuristicsVsPlanning} \cite{Dor1999SOKOBANAO} \cite{Froleyks2017UsingAA} \cite{SokoRelevanceCuts}
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.2]{CrookedLevel}
+\caption{A small (without the light gray part) and a large level (with the light gray part) with non-rectangular Structure}
+\label{fig:levelCrooked}
+\end{figure}
 \section{Encodings}
@@
 Those are used for simplifying complex rules; the first rule without the aggregates can be seen in listing \ref{list:BadFrame}.
-\lstinputlisting{../EncodingsClean/BadSplitFrameAxiom.lp}
+\lstinputlisting{../EncodingsClean/BadSplitFrameAxiom.txt}
 \begin{lstlisting}[caption={Frame Axiom X},label=list:BadFrame]
@@
 
-\subsection{Optimizations}
+\subsection{Strategies}
 
@@
 There are three different approaches I tried to remove deadlocks.
-The Forbidden Fields optimization (listing \ref{list:Forbidden}) works by generating a list of forbidden fields.
+The Forbidden Fields strategy (listing \ref{list:Forbidden}) works by generating a list of forbidden fields.
 Those fields are creating deadlocks when the player pushes a box onto them.
@@
 This approach can be seen in listing \ref{list:WBT}.
-This optimization works best in rectangular levels by simplifying the forbidden check for this special case.
+This strategy works best in rectangular levels by simplifying the forbidden check for this special case.
 The implementaion does not look at tunnels or similar structures but only checks for the outermost lines and rows.
@@
 This is a 2x2 square where each field is either a box or a wall, which would imply that all those boxes would have to stay in their places.
-In line 8 of the optimization we can see that the box D does not have a target underneath, which qualifies the situation to be a deadlock.
+In line 8 of the strategy implementation we can see that the box D does not have a target underneath, which qualifies the situation to be a deadlock.
 This type of deadlock happens a lot in crammed levels with a high box to field ratio.
@@
 
-In the following I will show two more optimizations working with move evaluation. This means eliminating movies that will not be relevant for a shortest plan.
+In the following I will show two more strategies working with move evaluation. This means eliminating moves that will not be relevant for a shortest plan.
 
@@
 
-The optimization in listing \ref{list:MOOB} is very simple but can remove a lot of moves from grounding, which would be removed after ramifications when checking integrity constraints otherwise.
+The strategy in listing \ref{list:MOOB} is very simple but can remove a lot of moves from grounding, which would be removed after ramifications when checking integrity constraints otherwise.
 This is more important in levels with a high horizon constant, i.e.
 a long shortest plan.
-The optimization works by preventing all moves that would lead the player of the playable area, for instance a move to the left, while the player is at the leftmost position of the level.
+The strategy works by preventing all moves that would lead the player of the playable area, for instance a move to the left, while the player is at the leftmost position of the level.
 
@@
 This would imply that the player moved back and forth without actually changing the gamestate, which means the plan is not relevant for finding the shortest plan in the level instance.
-
-\section{Evaluation}
-
-\section{Experiments and Results}
-
-\chapter{Discussion of Results}
-compare chapter State of the Art with chapter My Work \cite{LifschitzASP}
+\chapter{Evaluation}
+\chapter{Discussion}
+compare chapter Related Work with chapter My Work \cite{LifschitzASP}
 

### 2018-10-02 ResultParser.py

@@ -1,0 +1,123 @@
+import os, xlwt, string
+
+book = xlwt.Workbook(encoding="utf-8")
+
+sheet1 = book.add_sheet("Sheet 1", cell_overwrite_ok=True)
+
+index = 0
+sheet1.write(index, 0, "ENCODING")
+sheet1.write(index, 1, "OPTIMIZATIONS")
+sheet1.write(index, 2, "SIZE")
+sheet1.write(index, 3, "BOXRATIO")
+sheet1.write(index, 4, "STRUCTURE")
+sheet1.write(index, 5, "ASSIGNMENTS")
+sheet1.write(index, 6, "ZONES")
+sheet1.write(index, 7, "RESULT")
+sheet1.write(index, 8, "CHOICES")
+sheet1.write(index, 9, "CONFLICTS")
+sheet1.write(index, 10, "RESTARTS")
+sheet1.write(index, 11, "ATOMS")
+sheet1.write(index, 12, "RULES")
+sheet1.write(index, 13, "VARIABLES")
+sheet1.write(index, 14, "CONSTRAINTS")
+sheet1.write(index, 15, "TIME")
+index += 1
+
+for file in os.listdir("Results"):
+	
+	myfile = open(os.path.join("Results",file),"r")
+	linelist = myfile.readlines()
+	myname = os.path.basename(myfile.name)
+	thename = myname
+	myname = myname[7:]
+	if "Split" in myname:
+		sheet1.write(index, 0, "Split")
+		myname = myname[4:]
+	else:
+		sheet1.write(index, 0, "Plain")
+	if myname[:1] == "C":
+		sheet1.write(index, 1, "Clutter")
+		myname = myname[1:]
+	elif myname[:2] == "FF":
+		sheet1.write(index, 1, "ForbiddenFields")
+		myname = myname[2:]
+	elif myname[:4] == "MOOB":
+		sheet1.write(index, 1, "MovesOutOfBounds")
+		myname = myname[4:]
+	elif myname[:3] == "NTB":
+		sheet1.write(index, 1, "NoTakebacks")
+		myname = myname[3:]
+	elif myname[:4] == "WBTC":
+		sheet1.write(index, 1, "WallBoxTargetCheck")
+		myname = myname[4:]
+	elif myname[:5] == "Type1":
+		sheet1.write(index, 1, "Clutter, MOOB, NTB")
+		myname = myname[5:]
+	elif myname[:5] == "Type2":
+		sheet1.write(index, 1, "Clutter, MOOB, NTB, WBTC")
[…]

### 2018-10-02 Stuff/Parser.py

@@ -1,5 +1,5 @@
 def read_map():
-    f = open('InstancesClean/Inst-A-Z20-4Halle.lp', 'r')
+    f = open('Stuff/FirstLevel.lp', 'r')
     sr = f.read()
     # f = open('atlist','r')
 

### 2018-10-02 Stuff/Visualizer.py

@@ -16,7 +16,7 @@ class Visualizer(QtWidgets.QApplication):
         my_grid = QtWidgets.QGridLayout()
         w.setLayout(my_grid)
 
-        f = open('InstancesClean/Inst-A-Z20-4Halle.lp', 'r')
+        f = open('Stuff/FirstLevel.lp', 'r')
         sr = f.read()
         n = get_max(sr)
         self.redraw(my_grid, w)

### 2018-10-02 bashscript.sh

@@ -1,7 +1,7 @@
 #!/bin/bash
 plain="EncodingsClean/SokobanPlainEncoding.lp"
 inc="EncodingsClean/SokobanIncEncoding.lp"
-split="EncodingsClean/SokobanSplitEncoding.lp"
+split="EncodingsClean/SokobanSplitEncodingB.lp"
 splitinc="EncodingsClean/SokobanSplitIncEncoding.lp"
 plainass="EncodingsClean/Assignments/SokobanPlainEncodingAssignments.lp"
 incass="EncodingsClean/Assignments/SokobanIncEncodingAssignments.lp"
@@ -14,31 +14,33 @@ moob="EncodingsClean/Optimizations/MovesOutOfBounds.optimization.lp"
 ntb="EncodingsClean/Optimizations/NoTakebacks.optimization.lp"
 wbtc="EncodingsClean/Optimizations/WallBoxTargetCheck.optimization.lp"
 
-spclutter="EncodingsClean/Optimizations Split/SplitClutter.optimization.lp"
-spforfie="EncodingsClean/Optimizations Split/SplitForbiddenFields.optimization.lp"
-spmoob="EncodingsClean/Optimizations Split/SplitMovesOutOfBounds.optimization.lp"
-spntb="EncodingsClean/Optimizations Split/SplitNoTakebacks.optimization.lp"
-spwbtc="EncodingsClean/Optimizations Split/SplitWallBoxTargetCheck.optimization.lp"
+spclutter="EncodingsClean/OptimizationsSplit/SplitClutter.optimization.lp"
+spforfie="EncodingsClean/OptimizationsSplit/SplitForbiddenFields.optimization.lp"
+spmoob="EncodingsClean/OptimizationsSplit/SplitMovesOutOfBounds.optimization.lp"
+spntb="EncodingsClean/OptimizationsSplit/SplitNoTakebacks.optimization.lp"
+spwbtc="EncodingsClean/OptimizationsSplit/SplitWallBoxTargetCheck.optimization.lp"
 
 index=0
 
 for inst in "InstancesClean"/Inst-A-Z*.lp
 do
-	index=$(($index+1))
-	clingo $plain $inst $clutter --stats -q2 --time-limit=900 >> resultsC${index}.txt
-	clingo $plain $inst $forfie --stats -q2 --time-limit=900 >> resultsFF${index}.txt
-	clingo $plain $inst $moob --stats -q2 --time-limit=900 >> resultsMOOB${index}.txt
[…]

### 2018-10-03 BATexFiles/Draft.tex
NEUE VERSION Draft.pdf ist viel weiter

@@
-\documentclass{report}
+\documentclass[12pt,a4paper,oneside]{report}
 \title{Efficient Planning for Sokoban with Answer Set Programming - A Case Study}
 \author{Thomas Verweyen}
-\date{\today}
+\date{30.09.2018}
+\usepackage[utf8]{inputenc}
+\usepackage[ngerman]{babel}
+\usepackage[T1]{fontenc}
 \usepackage{listings}
+\lstdefinelanguage{clingo}{
+  keywordstyle=[1]\usefont{OT1}{cmtt}{m}{n},%
+  keywordstyle=[2]\textbf,%
+  keywordstyle=[3]\usefont{OT1}{cmtt}{m}{n},%\textit
+  alsoletter={\#,\&},% 
+keywords=[1]{not,from,import,def,if,else,return,while,break,and,or,for,in,del,and,class},% 
+keywords=[2]{\#const,\#show,\#minimize,\#base,\#theory,\#count,\#external,\#program,\#script,\#end,\#heuristic,\#edge,\#project,\#show},%
+  keywords=[3]{&,&dom,&sum,&diff,&show,&minimize},%
+  morecomment=[l]{\#\ },%
+  morecomment=[l]{\%\ },%
+  commentstyle={\color{darkgray}}%
+}
+\lstset{numbers=left,numberblanklines=false,basicstyle=\ttfamily\footnotesize,language=clingo}
+\usepackage{multirow}
 \usepackage{graphicx}
 \usepackage{titlesec}
+\usepackage[top=1.25in, bottom=1in, left=4cm, right=3.1cm]{geometry}
 \graphicspath{{./images/}}
@@
 \maketitle
-\renewcommand*\contentsname{Summary}
+\renewcommand*\contentsname{Table of Contents}
 \tableofcontents
@@
 The question whether Sokoban is PSPACE-complete remained open until 1997, when Joseph C.
-Culberson published his paper "Sokoban is PSPACE-complete" \cite{PSpaceComplete}.
+Culberson published his paper 'Sokoban is PSPACE-complete' \cite{PSpaceComplete}.
 He found a transformation algorithm to transform an LBA into a level of Sokoban, with the solution to the level having a number of steps in $\Theta (|w|+t(|w|))$, where $w$ is the input and $t(|w|)$ is the amount of transitions the LBA made.
@@
 The first type is a rectangular level with the X and Y dimension differing by 1.
-The second type is created by connecting 2x2 fields at their edges, see \ref{fig:levelCrooked}.
+The second type is created by connecting $2 \times 2$ fields at their edges, see \ref{fig:levelCrooked}.
 \begin{figure}[ht]
@@
 A choice rule picks the moves (l.6) and the succeeding states are derived from those by using effect and frame axioms (ll.8-11).
-In the first few lines basic information is created, namely a $time/1$ predicate for each timestep (l.1), a $dir/2$ predicate for each of the four possible directions to move in (l.3) and an $at/4$ predicate to set the initial positions of objects as their position at time T=0.
+In the first few lines basic information is created, namely a $time/1$ predicate for each timestep (l.1), a $dir/2$ predicate for each of the four possible directions to move in (l.3) and an $at/4$ predicate to set the initial positions of objects as their position at time $T=0$.
 Additionally the encoding needs integrity constraints to ensure that the plans found are legal.
@@
 To achieve an encoding that is able to, I will take advantage of incremental solving.
+\\
 \subsection{Incremental Encoding} \label{IncEnc}
@@
 In the incremental encoding I carry the time in the $box/3$ prediate to ensure that the constraint in line 21 only affects the target at time t.
-
+\\
 \subsection{Split Encoding}
@@
 The split encoding uses a technique common in ASP planning.
-When all coordinates are split into a vertical and a horizontal information the grounder does not have to ground all combinations of coordinates ($X * Y$) but only the sum of coordinates ($X + Y$).
+When all coordinates are split into a vertical and a horizontal information the grounder does not have to ground all combinations of coordinates ($X \cdot Y$) but only the sum of coordinates ($X + Y$).
 This has the advantage of less ground instances, but the disadvantage of a harder solving process.
@@
 
-My third approach to deadlock detection (see listing \ref{list:Clutter}) recognizes a "clutter" of boxes (see Fig.
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.2]{deadlocktypes}
+\caption{'Clutter' in four configurations}
+\label{fig:clutter}
+\end{figure}
+
+My third approach to deadlock detection (see listing \ref{list:Clutter}) recognizes a 'clutter' of boxes (see Fig.
 \ref{fig:clutter}).
-This is a 2x2 square where each field is either a box or a wall, which would imply that all those boxes would have to stay in their places.
+This is a $2 \times 2$ square where each field is either a box or a wall, which would imply that all those boxes would have to stay in their places.
 In line 8 of the strategy implementation we can see that the box D does not have a target underneath, which qualifies the situation to be a deadlock.
@@
 Those very common deadlocks are very important to find and remove to reduce the search tree in those levels efficiently.
-
+\\
 \subsubsection{Move Evaluation}
@@
 The last move evaluation seen in listing \ref{list:NoTake} tries to recognize moves that have no effect.
-This is done by checking for move instances $move(DX,DY,T)$ and $move(-DX,-DY,T-1)$ where no box was moved in timestep T-1.
+This is done by checking for move instances $move(DX,DY,T)$ and $move(-DX,-DY,T-1)$ where no box was moved in timestep $T-1$.
 This would imply that the player moved back and forth without actually changing the gamestate, which means the plan is not relevant for finding the shortest plan in the level instance.
 \chapter{Evaluation}
+\begin{table}[]
+\begin{tabular}{l|l|l|l|l|l|l|}
+\cline{2-7}
+                                  & \multicolumn{6}{l|}{Rectangular}                              \\ \cline{2-7} 
+                                  & \multicolumn{3}{l|}{Small} & \multicolumn{3}{l|}{Large} \\ \hline
+\multicolumn{1}{|l|}{\# of Boxes} & 4      & 5       & 6       & 6          & 8     & 10    \\ \hline
+\multicolumn{1}{|l|}{Basic}       & 2.97   & 34.157  & 10.615  & 46.283     & 900   & 900   \\ \hline
+\multicolumn{1}{|l|}{Clutter}     & 0.473  & 56.87   & 5.067   & 10.528     & 900   & 900   \\ \hline
+\multicolumn{1}{|l|}{FF}          & 2.946  & 29.984  & 9.566   & 40.181     & 900   & 900   \\ \hline
+\multicolumn{1}{|l|}{MOOB}        & 1.623  & 21.301  & 6.717   & 139.57     & 900   & 900   \\ \hline
+\multicolumn{1}{|l|}{NTB}         & 1.051  & 57.824  & 14.575  & 151.681    & 900   & 900   \\ \hline
+\multicolumn{1}{|l|}{WBTC}        & 2.655  & 30.85   & 9.611   & 42.185     & 900   & 900   \\ \hline
+\multicolumn{1}{|l|}{Type1}       & 0.525  & 24.504  & 4.483   & 76.499     & 900   & 900   \\ \hline
+\multicolumn{1}{|l|}{Type2}       & 0.523  & 24.728  & 4.483   & 78.336     & 900   & 900   \\ \hline
+\end{tabular}
+\end{table}
+\begin{table}[]
+\begin{tabular}{l|l|l|l|l|l|l|}
+\cline{2-7}
+                                  & \multicolumn{6}{l|}{Non-Rectangular}                             \\ \cline{2-7} 
+                                  & \multicolumn{3}{l|}{Small} & \multicolumn{3}{l|}{Large} \\ \hline
+\multicolumn{1}{|l|}{\# of Boxes} & 4       & 5       & 6      & 6       & 8       & 10     \\ \hline
+\multicolumn{1}{|l|}{Basic}       & 2.648   & 8.132   & 2.655  & 900     & 900     & 900    \\ \hline
+\multicolumn{1}{|l|}{Clutter}     & 1.1     & 3.107   & 2.127  & 900     & 900     & 900    \\ \hline
+\multicolumn{1}{|l|}{FF}          & 2.234   & 8.054   & 2.473  & 900     & 900     & 900    \\ \hline
+\multicolumn{1}{|l|}{MOOB}        & 2.23    & 7.953   & 2.466  & 900     & 900     & 900    \\ \hline
+\multicolumn{1}{|l|}{NTB}         & 2.132   & 5.635   & 1.914  & 900     & 900     & 900    \\ \hline
+\multicolumn{1}{|l|}{WBTC}        & 2.166   & 8.167   & 2.423  & 900     & 900     & 900    \\ \hline
+\multicolumn{1}{|l|}{Type1}       & 1.496   & 4.736   & 2.285  & 900     & 900     & 900    \\ \hline
+\multicolumn{1}{|l|}{Type2}       & 1.513   & 4.739   & 2.264  & 900     & 900     & 900    \\ \hline
+\end{tabular}
+\end{table}
+\begin{table}[]
+\begin{tabular}{l|l|l|l|l|l|}
+\cline{2-6}
+                              & \multicolumn{3}{l|}{Rectangular}         & \multicolumn{2}{l|}{Non-Rectangular} \\ \cline{2-6} 
+                              & \multicolumn{2}{l|}{Small} & Large & \multicolumn{2}{l|}{Small}       \\ \hline
+\multicolumn{1}{|l|}{\# of Boxes} & 5            & 8           & 10    & 4            & 5            \\ \hline
+\multicolumn{1}{|l|}{Basic}   & 535.227      & 900         & 900   & 54.624       & 57.636       \\ \hline
+\multicolumn{1}{|l|}{Clutter} & 193.626      & 900.031     & 900   & 26.104       & 26.351       \\ \hline
+\multicolumn{1}{|l|}{FF}      & 586.862      & 900.007     & 900   & 51.58        & 52.374       \\ \hline
+\multicolumn{1}{|l|}{MOOB}    & 305.945      & 900.013     & 900   & 51.586       & 52.889       \\ \hline
+\multicolumn{1}{|l|}{NTB}     & 309.895      & 900.009     & 900   & 110.652      & 121.758      \\ \hline
+\multicolumn{1}{|l|}{WBTC}    & 605.187      & 900.011     & 900   & 52.459       & 53.125       \\ \hline
+\multicolumn{1}{|l|}{Type1}   & 38.671       & 534.863     & 900   & 35.324       & 35.413       \\ \hline
+\multicolumn{1}{|l|}{Type2}   & 39.421       & 536.912     & 900   & 35.213       & 35.237       \\ \hline
+\end{tabular}
+\end{table}
 \chapter{Discussion}
@@
 
+\newpage
+\chapter*{Selbstständigkeitserklärung}
+\newpage
+\chapter*{Danksagung}
+
 \end{document}

### 2018-10-03 ResultParser.py

@@ -28,13 +28,13 @@ for file in os.listdir("Results"):
 	myfile = open(os.path.join("Results",file),"r")
 	linelist = myfile.readlines()
 	myname = os.path.basename(myfile.name)
+	print myname
 	thename = myname
-	myname = myname[7:]
-	if "Split" in myname:
-		sheet1.write(index, 0, "Split")
-		myname = myname[4:]
-	else:
-		sheet1.write(index, 0, "Plain")
+	print "heyo"
+	print myname
+	while myname[:1] in "resultsSplitInst":
+		myname = myname[1:]
+	print myname
 	if myname[:1] == "C":
 		sheet1.write(index, 1, "Clutter")
 		myname = myname[1:]
@@ -94,23 +94,36 @@ for file in os.listdir("Results"):
 		sheet1.write(index, 4, "Winkel")
 	all = string.maketrans('','')
 	nodigs = all.translate(all, string.digits)
-	if len(linelist) > 28:
-		if "UNKNOWN" not in myfile.read():
-			print thename
-			j = 0
-			while "Calls" not in linelist[j]:
-				j += 1
-			j+=1
-			i = j+3
-			sheet1.write(index, 15, linelist[j].split(' : ')[1].split(' ')[0])
-			sheet1.write(index, 8, linelist[i].split(' : ')[1].split(' ')[0].translate(all,nodigs))
-			sheet1.write(index, 9, linelist[i+1].split(' : ')[1].split(' ')[0].translate(all,nodigs))
-			sheet1.write(index, 10, linelist[i+2].split(' : ')[1].split(' ')[0].translate(all,nodigs))
-			sheet1.write(index, 11, linelist[i+12].split(' : ')[1].split(' ')[0].translate(all,nodigs))
-			sheet1.write(index, 12, linelist[i+13].split(' : ')[1].split(' ')[0].translate(all,nodigs))
-			sheet1.write(index, 13, linelist[i+14].split(' : ')[1].split(' ')[0].translate(all,nodigs))
-			sheet1.write(index, 14, linelist[i+15].split(' : ')[1].split(' ')[0].translate(all,nodigs))
-		if "TIME LIMIT" in myfile.read():
+	if len(linelist) > 10:
+		print thename
+		j = 0
+		while "Calls" not in linelist[j]:
+			j += 1
+		j+=1
+		sheet1.write(index, 15, linelist[j].split(' : ')[1].split(' ')[0])
[…]

### 2018-10-03 Stuff/Parser.py

@@ -1,5 +1,5 @@
 def read_map():
-    f = open('Stuff/FirstLevel.lp', 'r')
+    f = open('InstancesClean/Inst-A-Z32-10Winkel.lp', 'r')
     sr = f.read()
     # f = open('atlist','r')
 

### 2018-10-03 Stuff/Visualizer.py

@@ -16,7 +16,7 @@ class Visualizer(QtWidgets.QApplication):
         my_grid = QtWidgets.QGridLayout()
         w.setLayout(my_grid)
 
-        f = open('Stuff/FirstLevel.lp', 'r')
+        f = open('InstancesClean/Inst-A-Z32-10Winkel.lp', 'r')
         sr = f.read()
         n = get_max(sr)
         self.redraw(my_grid, w)
